---
title: "data_process"
author: "Yuna"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: '2025-01-07'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,message =  F,warning=F,results = "hide")
```

### Working path

```{r}
setwd("~/shortcut2project/simulation/")
```

### Library packages

```{r}
library(DBI)
library(RSQLite)
library(dplyr)
library(ggplot2)
library(latex2exp)
```

### Create a database

```{r results='hold'}
singleTrait_db<-dbConnect(RSQLite::SQLite(),dbname="~/shortcut2project/simulation/singleTrait.sqlite")
on.exit(dbDisconnect(singleTrait_db))

dbListTables(singleTrait_db)
```

### Simulation paramters
```{r}
N=10000
```

# Data storage to the database

## save simulation results of the burn-in phase (-10 - 0 N generations)
```{r}
paras<-data.frame(Vs=rep(c(1,10,100),each=3),sdMutation=rep(c(0.05,0.1,0.5),3))

for(npara in 1:nrow(paras)){
  Vs=paras[npara,"Vs"]
  sdMutation=paras[npara,"sdMutation"]
  
  path=paste0("Vs",Vs,"sd",sdMutation,"/burn_in/")
    
    for(i in 1:100){
      stats<-read.table(paste0(path,"stats_",i,".txt"),header=F,col.names=c("replicate","generation","meanPheno","varPheno","meanFitness","varFitness"))
      stats<-mutate(stats,Vs=Vs,sdMutation=sdMutation)
      dbWriteTable(singleTrait_db,"burnin_stats",stats,append=T)
      
      # information of segregating sites
       muta<-read.table(paste0(path,"muta_",i,".txt"),
                     col.names=c("replicate","generation","id","position","frequency","effectSize","aroseTick"))
       muta<-mutate(muta,Vs=Vs,sdMutation=sdMutation)
       dbWriteTable(singleTrait_db,name="burnin_muta",muta,append=T)
    }
}
```

### Calculate the mean genetic variance of the last 1000 generations as a base of gradual shift speeds
```{r}
burnin_stats<-tbl(singleTrait_db,"burnin_stats")

burnin_stats2_end<-filter(burnin_stats,generation>99000)%>%
  group_by(Vs,sdMutation,generation)%>%
  summarise(varPheno_mean=mean(varPheno))
parameters<-group_by(burnin_stats2_end,Vs,sdMutation)%>%
  summarise(SSE_VG=mean(varPheno_mean))%>%as.data.frame
dbWriteTable(singleTrait_db,"parameters",parameters)

# save for writing in gradual_shift.sh
write.table(parameters$Vs,file="Vs_ls.txt",col.names = F,row.names = F)
write.table(parameters$sdMutation,file="sd_ls.txt",col.names = F,row.names = F)
write.table(parameters$SSE_VG,file="VG_ls.txt",col.names = F,row.names = F)

parameters
```

### Calculate the fixed effect at the end of the stabilizing-selection phase
These are used to add in the phenotypic values in the next simulation steps, since fixations information were not recorded in the population files.
```{r}
for(Vs in c(1,10,100)){
  for(sdMutation in c(0.05,0.1,0.5)){
    
    SSE_fixed100<-c()
    
    for(replicate in 1:100){
      fixedMuta<-read.table(paste0('Vs',Vs,'sd',sdMutation,'/burn_in/fixedMuta_',replicate,'.txt'),
                            col.names=c('replicate','id','position','effectSize','aroseTick','endTick'))
      SSE_fixed100<-c(SSE_fixed100,sum(fixedMuta$effectSize))
    }
    
    write.table(SSE_fixed100,file=paste0('Vs',Vs,'sd',sdMutation,'/SSE_fixed.txt'),col.names = F,row.names = F)
  }
}
```

## save simulation results of the moving-optimum phase (0 - 10 N generations)
```{r}
# 175 speeds
paras<-rbind(data.frame(Vs=1,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.35,0.38,0.39,0.4,0.5,0.53,0.55,0.8,1.0,1.2,1.4,1.5,1.6)),
             data.frame(Vs=1,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.0,1.1,1.2,1.3,1.6,1.65,2.2,2.8,2.9,3.0,3.2,3.4,3.6)),
             data.frame(Vs=1,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.0,2.0,3.0,6.5,10.0,11.0,12.0,12.5,13.5,14.0,15.0,16.0,17.0,18.0,18.2,20.0,20.5,21.0)),
             data.frame(Vs=10,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.07,0.08,0.09,0.1,0.11,0.2,0.3,0.4)),
             data.frame(Vs=10,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.15,0.18,0.2,0.22,0.23,0.24,0.3,0.4,0.5,0.6,0.7,0.8)),
             data.frame(Vs=10,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.3,2.3,2.32,2.35,2.4,3.3,3.35,4.3,1.0,4.5,4.6,4.9,5.0,5.2,5.4,6.0)),
             data.frame(Vs=100,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.025,0.03,0.031,0.035,0.04,0.05,0.1,0.2)),
             data.frame(Vs=100,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.03,0.04,0.045,0.05,0.055,0.057,0.1,0.2,0.3)),
             data.frame(Vs=100,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.35,0.38,0.39,0.4,0.5,0.52,0.55,0.7,0.9,1.1,1.3,1.4,1.5))
)

for(npara in 1:nrow(paras)){
  Vs=paras[npara,"Vs"]
  sdMutation=paras[npara,"sdMutation"]
  speedCoef=paras[npara,"speedCoef"]
  
  path=paste0("Vs",Vs,"sd",sdMutation,"/gradual/")
  
  for(i in 1:100){
    stats<-read.table(paste0(path,"stats_",speedCoef,"_",i,".txt"),header=F,col.names=c("replicate","generation","meanPheno","varPheno","meanFitness","varFitness"))
    stats<-mutate(stats,Vs=Vs,sdMutation=sdMutation,speedCoef=speedCoef)
    dbWriteTable(singleTrait_db,name="gradual_stats",stats,append=T)
    
    # information of segregating sites
    muta<-read.table(paste0(path,"muta_",speedCoef,"_",i,".txt"),
                     col.names=c("replicate","generation","id","position","frequency","effectSize","aroseTick"))
    muta<-mutate(muta,Vs=Vs,sdMutation=sdMutation,speedCoef=speedCoef)
    dbWriteTable(singleTrait_db,name="gradual_muta",muta,append=T)
  }
}
```


# Data process

## Phenotype

### Phenotypic changes over time
burn_in: calculate the mean and sd of 100 replicates at each generation
```{r}
burnin_stats<-tbl(singleTrait_db,"burnin_stats")
burnin_stats2<-filter(burnin_stats, generation%%1000==1 | generation>=99000)%>%
  group_by(Vs,sdMutation,generation)%>%
  summarise(
    varPheno_mean=mean(varPheno),
    varPheno_sd=sd(varPheno)
  )
burnin_stats2<-as.data.frame(burnin_stats2)
```

gradual: calculate the mean and sd of 100 replicates at each generation
``` {r}
gradual_stats<-tbl(singleTrait_db,"gradual_stats")

gradual_stats2<-filter(gradual_stats, generation<=101000 | generation%%200==0 | generation >199000)%>%
  group_by(Vs,sdMutation,speedCoef,generation)%>%
  summarise(meanPheno_mean=mean(meanPheno),
            meanPheno_sd=sd(meanPheno),
            varPheno_mean=mean(varPheno),
            varPheno_sd=sd(varPheno),
            meanFitness_mean=mean(meanFitness),
            meanFitness_sd=sd(meanFitness),
            varFitness_mean=mean(varFitness),
            varFitness_sd=sd(varFitness)
  )
parameters<-tbl(singleTrait_db,"parameters")
gradual_stats2<-left_join(gradual_stats2,parameters,by=c("Vs","sdMutation"))

gradual_stats2<-as.data.frame(gradual_stats2)
gradual_stats2<-mutate(gradual_stats2,optimum=speedCoef*sqrt(SSE_VG)*(generation-10*N),
                       gap=optimum-meanPheno_mean,
                       speed=speedCoef * sqrt(SSE_VG),
                       sigmaOmega2=sdMutation^2/Vs)

so=unique(gradual_stats2$sigmaOmega2)%>%sort
so=as.character(format(so[-5],scientific = T))
gradual_stats2$sigmaOmega2<-factor(format(gradual_stats2$sigmaOmega2,scientific=T),level=so,label=so)
```

extincted: calculate the mean and sd of 100 replicates at each generation
```{r}
paras<-data.frame(Vs=c(1,1,1,10,10,10,100,100,100),
                  sdMutation=c(0.05,0.1,0.5,0.05,0.1,0.5,0.05,0.1,0.5),
                  speedCoef=c(1.7,4,22,0.6,1,7,0.4,0.5,1.7))
extincted_stats<-c()
for(npara in 1:nrow(paras)){
  Vs=paras[npara,"Vs"]
  sdMutation=paras[npara,"sdMutation"]
  speedCoef=paras[npara,"speedCoef"]
  
  path=paste0("Vs",Vs,"sd",sdMutation,"/gradual/")
  
  for(i in 1:100){
    stats<-read.table(paste0(path,"stats_",speedCoef,"_",i,".txt"),header=F,col.names=c("replicate","generation","meanPheno","varPheno","meanFitness","varFitness"))
    stats<-mutate(stats,Vs=Vs,sdMutation=sdMutation,speedCoef=speedCoef)
    extincted_stats<-rbind(extincted_stats,stats)
  }
}

extincted_stats2<-filter(extincted_stats, generation<=101000 | generation%%100==0)%>%
  group_by(Vs,sdMutation,speedCoef,generation)%>%
  summarise(n_replicates=n(),meanPheno_mean=mean(meanPheno),
            meanPheno_sd=sd(meanPheno),
            varPheno_mean=mean(varPheno),
            varPheno_sd=sd(varPheno),
            meanFitness_mean=mean(meanFitness),
            meanFitness_sd=sd(meanFitness),
            varFitness_mean=mean(varFitness),
            varFitness_sd=sd(varFitness)
  )

singleTrait_db<-dbConnect(RSQLite::SQLite(),dbname="~/shortcut/B6/simulation/singleTrait.sqlite")
parameters<-tbl(singleTrait_db,"parameters")%>%as.data.frame
extincted_stats2<-left_join(extincted_stats2,parameters,by=c("Vs","sdMutation"))

N=1e4
extincted_stats2<-mutate(extincted_stats2,optimum=speedCoef*sqrt(SSE_VG)*(generation-10*N),
                       gap=optimum-meanPheno_mean,
                       speed=speedCoef * sqrt(SSE_VG),
                       sigmaOmega2=sdMutation^2/Vs)

so=unique(extincted_stats2$sigmaOmega2)%>%sort
so=as.character(format(so,scientific = T))
extincted_stats2$sigmaOmega2<-factor(format(extincted_stats2$sigmaOmega2,scientific=T),level=so,label=so)

load("stats2.RData")
save(burnin_stats2,gradual_stats2,extincted_stats2,file='stats2.RData')
```


### Mean phenotype in the stationary state
calculate the mean of 100 replicates at each of the last 1000 generations in two phases, respectively
```{r}
# stabilizing-selection equilibrium
burnin_stats<-tbl(singleTrait_db,"burnin_stats")
parameters<-tbl(singleTrait_db,"parameters")
SSE_df2<-filter(burnin_stats,generation>99000)%>%
  group_by(Vs,sdMutation,generation)%>%
  summarise(meanPheno_mean=mean(meanPheno),
            varPheno_mean=mean(varPheno),
            meanFitness_mean=mean(meanFitness))
SSE_df2<-left_join(SSE_df2,parameters,by=c("Vs","sdMutation"))
SSE_df2<-as.data.frame(SSE_df2)
SSE_df2<-mutate(SSE_df2,optimum=0,
                gap=optimum-meanPheno_mean)

# moving-optimum equilibrium
gradual_stats<-tbl(singleTrait_db,"gradual_stats")
MOE_df2<-filter(gradual_stats,generation>199000)%>%
  group_by(Vs,sdMutation,speedCoef,generation)%>%
  summarise(meanPheno_mean=mean(meanPheno),
            varPheno_mean=mean(varPheno),
            meanFitness_mean=mean(meanFitness))
MOE_df2<-left_join(MOE_df2,parameters,by=c("Vs","sdMutation"))
MOE_df2<-as.data.frame(MOE_df2)
MOE_df2<-mutate(MOE_df2,optimum=speedCoef*sqrt(SSE_VG)*(generation-10*N),
            gap=optimum-meanPheno_mean)

parameters<-as.data.frame(parameters)

save(SSE_df2,MOE_df2,parameters,file="df2.RData")
```



## Fixations

### Fixations changes over time
pick few speeds for following analyses
```{r}
paras_quantile<-rbind(data.frame(Vs=1,sdMutation=0.05,speedCoef=c(0.002,0.2,0.5,0.8,1.2,1.6)),
                      data.frame(Vs=1,sdMutation=0.1,speedCoef=c(0.002,0.5,1.3,2.2,2.9,3.6)),
                      data.frame(Vs=1,sdMutation=0.5,speedCoef=c(0.002,3,6.5,12,16,21)),
                      data.frame(Vs=10,sdMutation=0.05,speedCoef=c(0.002,0.08,0.11,0.2,0.3,0.4)),
                      data.frame(Vs=10,sdMutation=0.1,speedCoef=c(0.002,0.15,0.3,0.5,0.6,0.8)),
                      data.frame(Vs=10,sdMutation=0.5,speedCoef=c(0.002,1,2.4,3.35,4.9,6)),
                      data.frame(Vs=100,sdMutation=0.05,speedCoef=c(0.002,0.02,0.035,0.05,0.1,0.2)),
                      data.frame(Vs=100,sdMutation=0.1,speedCoef=c(0.002,0.02,0.05,0.1,0.2,0.3)),
                      data.frame(Vs=100,sdMutation=0.5,speedCoef=c(0.002,0.3,0.55,0.9,1.3,1.5))
                      )
```

Scan the whole evolutionary time with a 100-generation window, take average of 100 replicates
```{r}
gran=100

mo_fixedMuta2<-c()
for(npara in 1:nrow(paras_quantile)){
  Vs=paras_quantile[npara,"Vs"]
  sdMutation=paras_quantile[npara,"sdMutation"]
  speedCoef=paras_quantile[npara,"speedCoef"]
  
  for(replicate in 1:100){
    err<-try(df<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/gradual/fixedMuta_",speedCoef,"_",replicate,".txt"),
                        colClasses=c("integer","NULL","NULL","numeric","integer","integer"),
                        col.names=c("replicate","id","position","effectSize","aroseTick","endTick")),silent=T)
    
    if("try-error" %in% class(err)){next}
    
    df<-mutate(df,fixedTime=ifelse(aroseTick>10*N,endTick-aroseTick,endTick-10*N))
    df$endTick_cut<-cut(df$endTick,breaks=seq(10*N,20*N,gran))
    df$endTick_cut<-as.numeric(df$endTick_cut)*gran
  
    df2<-group_by(df,endTick_cut)%>%
    summarise(n_fixed=n(),n_posFixed=sum(effectSize>0),
            mean_fixedtime=mean(fixedTime),mean_posFixedtime=mean(fixedTime[effectSize>0]),
            mean_negFixedtime=mean(fixedTime[effectSize<0]),
            mean_effect=mean(effectSize),mean_posEffect=mean(effectSize[effectSize>0]),
            mean_negEffect=mean(effectSize[effectSize<0]))
    
    df2<-mutate(df2,Vs=Vs,sdMutation=sdMutation,speedCoef=speedCoef,replicate=replicate)
    mo_fixedMuta2<-rbind(mo_fixedMuta2,df2)
  }
}

mo_fixedMuta3_n<-group_by(mo_fixedMuta2,Vs,sdMutation,speedCoef,endTick_cut)%>%summarise(n_fixed=sum(n_fixed)/gran/100,n_posFixed=sum(n_posFixed)/gran/100)
mo_fixedMuta3<-group_by(mo_fixedMuta2,Vs,sdMutation,speedCoef,endTick_cut)%>%summarise_at(vars(-replicate),mean,na.rm=T)
mo_fixedMuta3$n_fixed<-mo_fixedMuta3_n$n_fixed
mo_fixedMuta3$n_posFixed<-mo_fixedMuta3_n$n_posFixed
mo_fixedMuta3<-left_join(mo_fixedMuta3,as.data.frame(parameters),by=c('Vs','sdMutation'))
mo_fixedMuta3<-mutate(mo_fixedMuta3,speed=speedCoef*sqrt(SSE_VG))

save(mo_fixedMuta2,mo_fixedMuta3,file="mo_fixedMuta.RData")
```


### Fixations in the stationary state
Summarize the total number, mean and standard deviation of effect size and fixation time of fixations in the last 1000 generations. In addition, split the fixations by the direction of effect size and summarize.

burnin:
```{r}
sse_fixed<-data.frame(Vs=NA,sdMutation=NA,replicate=NA,
                      n_fixed=NA,n_posFixed=NA,
                      
                      mean_effect=NA,sd_effect=NA,
                      mean_posEffect=NA,sd_posEffect=NA,
                      mean_negEffect=NA,sd_negEffect=NA,
                      
                      mean_fixedtime=NA,sd_fixedtime=NA,
                      mean_posFixedtime=NA,sd_posFixedtime=NA,
                      mean_negFixedtime=NA,sd_negFixedtime=NA
)

i=1
for(Vs in c(1,10,100)){
  for(sdMutation in c(0.05,0.1,0.5)){
    for(replicate in 1:100){
      fixedMuta<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/burn_in/fixedMuta_",replicate,".txt"),
                            colClasses=c("integer","NULL","NULL","numeric","integer","integer"),
                            col.names=c("replicate","id","position","effectSize","aroseTick","endTick"))
      
      # stabilizing selection equilibrium: last 0.1N generation:
      fixedMuta<-subset(fixedMuta,endTick > 9.9*N)
      
      sse_fixed[i,"Vs"]<-Vs
      sse_fixed[i,"sdMutation"]<-sdMutation
      sse_fixed[i,"replicate"]<-replicate
      
      if(nrow(fixedMuta)==0){
        sse_fixed[i,"n_fixed"]<-0
        sse_fixed[i,"n_posFixed"]<-0
        
        sse_fixed[i,-c(1:5)]<-NaN
        
      }else{
        sse_fixed[i,"n_fixed"]<-nrow(fixedMuta)
        sse_fixed[i,"n_posFixed"]<-sum(fixedMuta$effectSize>0)
        
        sse_fixed[i,"mean_effect"]<-mean(fixedMuta$effectSize)
        sse_fixed[i,"sd_effect"]<-sd(fixedMuta$effectSize)
        sse_fixed[i,"mean_posEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize > 0])
        sse_fixed[i,"sd_posEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize > 0])
        sse_fixed[i,"mean_negEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize < 0])
        sse_fixed[i,"sd_negEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize < 0])
        
        fixedTime=fixedMuta$endTick-fixedMuta$aroseTick
        sse_fixed[i,"mean_fixedtime"]<-mean(fixedTime)
        sse_fixed[i,"sd_fixedtime"]<-sd(fixedTime)
        sse_fixed[i,"mean_posFixedtime"]<-mean(fixedTime[fixedMuta$effectSize > 0])
        sse_fixed[i,"sd_posFixedtime"]<-sd(fixedTime[fixedMuta$effectSize > 0])
        sse_fixed[i,"mean_negFixedtime"]<-mean(fixedTime[fixedMuta$effectSize < 0])
        sse_fixed[i,"sd_negFixedtime"]<-sd(fixedTime[fixedMuta$effectSize < 0])
      }
      
      i<-i+1
    }
  }
}
```

moving optimum:
```{r}
moe_fixed<-data.frame(Vs=NA,sdMutation=NA,speedCoef=NA,replicate=NA,
                     n_fixed=NA,n_burninArisen=NA,n_posFixed=NA,n_posBurnin=NA,
                     
                     mean_effect=NA,sd_effect=NA,
                     mean_burninEffect=NA,sd_burninEffect=NA,
                     mean_shiftEffect=NA,sd_shiftEffect=NA,
                     mean_posEffect=NA,sd_posEffect=NA,
                     mean_negEffect=NA,sd_negEffect=NA,
                     mean_posBurninEffect=NA,sd_posBurninEffect=NA,
                     mean_posShiftEffect=NA,sd_posShiftEffect=NA,
                     mean_negBurninEffect=NA,sd_negBurninEffect=NA,
                     mean_negShiftEffect=NA,sd_negShiftEffect=NA,
                     
                     mean_fixedtime=NA,sd_fixedtime=NA,
                     mean_burninFixedtime=NA,sd_burninFixedtime=NA,
                     mean_shiftFixedtime=NA,sd_shiftFixedtime=NA,
                     mean_posFixedtime=NA,sd_posFixedtime=NA,
                     mean_negFixedtime=NA,sd_negFixedtime=NA,
                     mean_posBurninFixedtime=NA,sd_posBurninFixedtime=NA,
                     mean_posShiftFixedtime=NA,sd_posShiftFixedtime=NA,
                     mean_negBurninFixedtime=NA,sd_negBurninFixedtime=NA,
                     mean_negShiftFixedtime=NA,sd_negShiftFixedtime=NA)
i=1
for(npara in 1:nrow(paras)){
  Vs=paras[npara,'Vs']
  sdMutation=paras[npara,'sdMutation']
  speedCoef=paras[npara,'speedCoef']
  
  for(replicate in 1:100){
     err<-try(fixedMuta<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/gradual/fixedMuta_",speedCoef,"_",replicate,".txt"),
                                   colClasses=c("integer","NULL","NULL","numeric","integer","integer"),
                                   col.names=c("replicate","id","position","effectSize","aroseTick","endTick")), silent = T)
    
    if("try-error" %in% class(err)){next}
    
    # moving optimum equilibrium: last 0.1N generation
    fixedMuta<-subset(fixedMuta,endTick > 19.9*N)
    
    moe_fixed[i,"Vs"]<-Vs
    moe_fixed[i,"sdMutation"]<-sdMutation
    moe_fixed[i,"speedCoef"]<-speedCoef
    moe_fixed[i,"replicate"]<-replicate
    
    if(nrow(fixedMuta)==0){
      moe_fixed[i,"n_fixed"]<-0
      moe_fixed[i,"n_burninArisen"]<-0
      moe_fixed[i,"n_posFixed"]<-0
      moe_fixed[i,"n_posBurnin"]<-0
      
      moe_fixed[i,-c(1:8)]<-NaN
      
    }else{
      moe_fixed[i,"n_fixed"]<-nrow(fixedMuta)
      moe_fixed[i,"n_burninArisen"]<-sum(fixedMuta$aroseTick<=10*N)
      moe_fixed[i,"n_posFixed"]<-sum(fixedMuta$effectSize>0)
      moe_fixed[i,"n_posBurnin"]<-sum(fixedMuta$effectSize>0 & fixedMuta$aroseTick<=10*N)
      
      moe_fixed[i,"mean_effect"]<-mean(fixedMuta$effectSize)
      moe_fixed[i,"sd_effect"]<-sd(fixedMuta$effectSize)
      moe_fixed[i,"mean_burninEffect"]<-mean(fixedMuta$effectSize[fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"sd_burninEffect"]<-sd(fixedMuta$effectSize[fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"mean_shiftEffect"]<-mean(fixedMuta$effectSize[fixedMuta$aroseTick>10*N])
      moe_fixed[i,"sd_shiftEffect"]<-sd(fixedMuta$effectSize[fixedMuta$aroseTick>10*N])
      moe_fixed[i,"mean_posEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize > 0])
      moe_fixed[i,"sd_posEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize > 0])
      moe_fixed[i,"mean_negEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize < 0])
      moe_fixed[i,"sd_negEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize < 0])
      moe_fixed[i,"mean_posBurninEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize > 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"sd_posBurninEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize > 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"mean_posShiftEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize > 0 & fixedMuta$aroseTick>10*N])
      moe_fixed[i,"sd_posShiftEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize > 0 & fixedMuta$aroseTick>10*N])
      moe_fixed[i,"mean_negBurninEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize < 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"sd_negBurninEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize < 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"mean_negShiftEffect"]<-mean(fixedMuta$effectSize[fixedMuta$effectSize < 0 & fixedMuta$aroseTick>10*N])
      moe_fixed[i,"sd_negShiftEffect"]<-sd(fixedMuta$effectSize[fixedMuta$effectSize < 0 & fixedMuta$aroseTick>10*N])
      
      fixedTime=fixedMuta$endTick-fixedMuta$aroseTick
      ind=which(fixedMuta$aroseTick<=10*N)
      fixedTime[ind]<-fixedMuta$endTick[ind] - 10*N
      moe_fixed[i,"mean_fixedtime"]<-mean(fixedTime)
      moe_fixed[i,"sd_fixedtime"]<-sd(fixedTime)
      moe_fixed[i,"mean_burninFixedtime"]<-mean(fixedTime[fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"sd_burninFixedtime"]<-sd(fixedTime[fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"mean_shiftFixedtime"]<-mean(fixedTime[fixedMuta$aroseTick>10*N])
      moe_fixed[i,"sd_shiftFixedtime"]<-sd(fixedTime[fixedMuta$aroseTick>10*N])
      moe_fixed[i,"mean_posFixedtime"]<-mean(fixedTime[fixedMuta$effectSize > 0])
      moe_fixed[i,"sd_posFixedtime"]<-sd(fixedTime[fixedMuta$effectSize > 0])
      moe_fixed[i,"mean_negFixedtime"]<-mean(fixedTime[fixedMuta$effectSize < 0])
      moe_fixed[i,"sd_negFixedtime"]<-sd(fixedTime[fixedMuta$effectSize < 0])
      moe_fixed[i,"mean_posBurninFixedtime"]<-mean(fixedTime[fixedMuta$effectSize > 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"sd_posBurninFixedtime"]<-sd(fixedTime[fixedMuta$effectSize > 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"mean_posShiftFixedtime"]<-mean(fixedTime[fixedMuta$effectSize > 0 & fixedMuta$aroseTick>10*N])
      moe_fixed[i,"sd_posShiftFixedtime"]<-sd(fixedTime[fixedMuta$effectSize > 0 & fixedMuta$aroseTick>10*N])
      moe_fixed[i,"mean_negBurninFixedtime"]<-mean(fixedTime[fixedMuta$effectSize < 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"sd_negBurninFixedtime"]<-sd(fixedTime[fixedMuta$effectSize < 0 & fixedMuta$aroseTick<=10*N])
      moe_fixed[i,"mean_negShiftFixedtime"]<-mean(fixedTime[fixedMuta$effectSize < 0 & fixedMuta$aroseTick>10*N])
      moe_fixed[i,"sd_negShiftFixedtime"]<-sd(fixedTime[fixedMuta$effectSize < 0 & fixedMuta$aroseTick>10*N])
    }
    
    rm(fixedMuta)
    i<-i+1
  }
}
```


## Segregating mutations
Collect the number of segregating mutations, mean effect and mean allelic frequency at every generation. In addition, collect by splitting the direction of effect

### Segregating mutations over time
```{r}
write.table(paras[,1],file="Vs_total.txt",row.names = F,col.names = F,quote=F)
write.table(paras[,2],file="sdMutation_total.txt",row.names = F,col.names = F,quote=F)
write.table(paras[,3],file="speed_total.txt",row.names = F,col.names = F,quote=F)
```

Analyzing through bash
```{bash}
sh muta_statistics.sh
```

burn in:
```{r}
muta2<-data.frame(Vs=NA,sdMutation=NA,speedCoef=NA,replicate=NA,generation=NA,
                  n_segregations=NA,mean_effect=NA,mean_frequency=NA,
                  n_posSegregations=NA,mean_posEffect=NA,mean_posFrequency=NA,
                  n_negSegregations=NA,mean_negEffect=NA,mean_negFrequency=NA)
for(Vs in c(1,10,100)){
  for(sdMutation in c(0.05,0.1,0.5)){
    muta2_df<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/burn_in/muta2.txt"),
                         col.names=c("speedCoef","replicate","generation","n_segregations","mean_effect","mean_frequency"))
    posMuta2_df<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/burn_in/posMuta2.txt"),
                            col.names=c("speedCoef","replicate","generation","n_posSegregations","mean_posEffect","mean_posFrequency"))
    negMuta2_df<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/burn_in/negMuta2.txt"),
                            col.names=c("speedCoef","replicate","generation","n_negSegregations","mean_negEffect","mean_negFrequency"))
    muta2_df_total<-left_join(muta2_df,posMuta2_df,by=c("speedCoef","replicate","generation"))%>%
      left_join(.,negMuta2_df,by=c("speedCoef","replicate","generation"))%>%
      mutate_if(is.numeric,coalesce,0)
    
    muta2_df_total<-data.frame(Vs=Vs,sdMutation=sdMutation,muta2_df_total)
    
    muta2<-rbind(muta2,muta2_df_total)
  }
}
muta2<-muta2[-1,]
burnin_muta2<-muta2
```

moving optimum:
```{r}
muta2<-data.frame(Vs=NA,sdMutation=NA,speedCoef=NA,replicate=NA,generation=NA,
                  n_segregations=NA,mean_effect=NA,mean_frequency=NA,
                  n_posSegregations=NA,mean_posEffect=NA,mean_posFrequency=NA,
                  n_negSegregations=NA,mean_negEffect=NA,mean_negFrequency=NA)
for(Vs in c(1,10,100)){
  for(sdMutation in c(0.05,0.1,0.5)){
    muta2_df<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/gradual/muta2.txt"),
                         col.names=c("speedCoef","replicate","generation","n_segregations","mean_effect","mean_frequency"))
    posMuta2_df<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/gradual/posMuta2.txt"),
                            col.names=c("speedCoef","replicate","generation","n_posSegregations","mean_posEffect","mean_posFrequency"))
    negMuta2_df<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/gradual/negMuta2.txt"),
                            col.names=c("speedCoef","replicate","generation","n_negSegregations","mean_negEffect","mean_negFrequency"))
    muta2_df_total<-left_join(muta2_df,posMuta2_df,by=c("speedCoef","replicate","generation"))%>%
      left_join(.,negMuta2_df,by=c("speedCoef","replicate","generation"))%>%
      mutate_if(is.numeric,coalesce,0)
    
    muta2_df_total<-data.frame(Vs=Vs,sdMutation=sdMutation,muta2_df_total)
    
    muta2<-rbind(muta2,muta2_df_total)
  }
}
muta2<-muta2[-1,]
shift_muta2<-muta2
```

Save the above results
```{r}
save(parameters,moe_fixed,sse_fixed,shift_muta2,burnin_muta2,file="muta2.RData")
```


### Calculate allele-frequency weighted mean and standard deviation
```{r}
library(data.table)

moe_muta_weightedMeanSd <- data.table()
for(npara in 1:nrow(paras)){
  Vs=paras[npara,"Vs"]
  sdMutation=paras[npara,"sdMutation"]
  speedCoef=paras[npara,"speedCoef"]
  
  path=paste0("Vs",Vs,"sd",sdMutation,"/gradual/")
  
  for(i in 1:100){
    err<-try(muta<-fread(paste0(path,"muta_",speedCoef,"_",i,".txt"),
                              colClasses=c("integer","integer","NULL","NULL","numeric","numeric","NULL"),
                              col.names=c("replicate","generation","frequency","effectSize")), silent = T)

    if("try-error" %in% class(err)){next}

    summ <- muta[, {
      sum_freq = sum(frequency)
      meanEffect = mean(effectSize)
      sdEffect = sd(effectSize)
      meanEffect_weighted = sum(frequency * effectSize) / sum_freq
      sdEffect_weighted = sqrt(sum(frequency * (effectSize - meanEffect_weighted)^2) / sum_freq)
      
      list(
        n_segre = .N,
        sum_frequency = sum_freq,
        mean_effect = meanEffect,
        sd_effect = sdEffect,
        mean_effect_weighted = meanEffect_weighted,
        sd_effect_weighted = sdEffect_weighted
      )
    }, by = .(generation)]

    summ[, `:=`(
      Vs = Vs,
      sdMutation = sdMutation,
      speedCoef = speedCoef,
      replicate = i
    )]
    
  moe_muta_weightedMeanSd <- rbindlist(list(moe_muta_weightedMeanSd, summ), use.names = TRUE, fill = TRUE)
  }
}

save(moe_muta_weightedMeanSd,file="moe_muta_weightedMeanSd.RData")
```

### Contribution of fixations
```{r}
fixed_contribution<-c()

for(npara in 73:nrow(paras)){
  Vs=paras[npara,"Vs"]
  sdMutation=paras[npara,"sdMutation"]
  speedCoef=paras[npara,"speedCoef"]
  
  for(replicate in 1:100){
    err<-try(fixedMuta<-read.table(paste0("Vs",Vs,"sd",sdMutation,"/gradual/fixedMuta_",speedCoef,"_",replicate,".txt"),colClasses=c("integer","NULL","NULL","numeric","integer","integer"),col.names=c("replicate","id","position","effectSize","aroseTick","endTick")), silent = T)

    if("try-error" %in% class(err)){next}
    
    fixed_df<-data.frame(Vs=Vs,sdMutation=sdMutation,speedCoef=speedCoef,replicate=replicate,
                         n_fixed=c(nrow(fixedMuta)),
                         n_posFixed=sum(fixedMuta$effectSize>0),
                         contribution_fixed=sum(fixedMuta$effectSize),
                         contribution_posFixed=sum(fixedMuta$effectSize[fixedMuta$effectSize>0]))
    
    fixed_contribution<-rbind(fixed_contribution,fixed_df)
  }
}

save(fixed_contribution,file="fixed_contribution.RData")
```


### disconnect the database
```{r}
dbDisconnect(singleTrait_db)
```

