---
title: "dynamics"
author: "Yuna"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2024-04-25"
---

### Working path

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/yuna/B6/simulation/",echo = T,message =  F,warning=F,results = "hide")
```

### Library packages

```{r}
library(DBI)
library(RSQLite)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggrepel)# annotation in figures
library(latex2exp)
library(scales)
```

### Connect the database

```{r results='hold'}
singleTrait_db<-dbConnect(RSQLite::SQLite(),dbname="~/shortcut/B6/simulation/singleTrait.sqlite")
# on.exit(dbDisconnect(singleTrait_db))

dbListTables(singleTrait_db)
```

### Simulation parameters

```{r}
N=10000
geno_len=999999+1
mutation_rate=7e-10
U=2*mutation_rate*geno_len
theta=14

paras<-rbind(data.frame(Vs=1,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.35,0.38,0.39,0.4,0.5,0.53,0.55,0.8,1.0,1.2,1.4,1.5,1.6)),
             data.frame(Vs=1,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.0,1.1,1.2,1.3,1.6,1.65,2.2,2.8,2.9,3.0,3.2,3.4,3.6)),
             data.frame(Vs=1,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.0,2.0,3.0,6.5,10.0,11.0,12.0,12.5,13.5,14.0,15.0,16.0,17.0,18.0,18.2,20.0,20.5,21.0)),
             data.frame(Vs=10,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.07,0.08,0.09,0.1,0.11,0.2,0.3,0.4)),
             data.frame(Vs=10,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.15,0.18,0.2,0.22,0.23,0.24,0.3,0.4,0.5,0.6,0.7,0.8)),
             data.frame(Vs=10,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.3,2.3,2.32,2.35,2.4,3.3,3.35,4.3,1.0,4.5,4.6,4.9,5.0,5.2,5.4,6.0)),
             data.frame(Vs=100,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.025,0.03,0.031,0.035,0.04,0.05,0.1,0.2)),
             data.frame(Vs=100,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.03,0.04,0.045,0.05,0.055,0.057,0.1,0.2,0.3)),
             data.frame(Vs=100,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.35,0.38,0.39,0.4,0.5,0.52,0.55,0.7,0.9,1.1,1.3,1.4,1.5))
)
```

### Global plotting settings

```{r}
theme_set(theme_classic())
label_sdMutation<-function(string){TeX(paste("$\\omega : $",string))}
```

# Genetic variance during the stabilizing-selection phase

```{r}
ggplot(burnin_stats2,aes(x=generation/N-10,y=varPheno_mean))+
  geom_line()+
  geom_ribbon(aes(ymin=varPheno_mean-varPheno_sd,ymax=varPheno_mean+varPheno_sd),col=NA,alpha=.3)+
  geom_hline(aes(yintercept=2*U*Vs),col='red')+ # estimates from House-of-Card model
  geom_hline(aes(yintercept=2*U*Vs/(1+Vs/(N*sdMutation^2))),col='red',linetype="dashed")+ # estimates from stochastic House-of-Card model
  facet_grid(Vs~sdMutation,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Genetic variance")
```

# Dynamics of phenotype during the moving-optimum phase

Pick out few speeds to plot
```{r}
load("data/processed/stats2.RData")

paras_quantile<-group_by(paras,Vs,sdMutation)%>%
  summarise(speed0=quantile(speedCoef,0),
            speed20=quantile(speedCoef,0.2),
            speed40=quantile(speedCoef,0.4),
            speed60=quantile(speedCoef,0.6),
            speed80=quantile(speedCoef,0.8),
            speed100=quantile(speedCoef,1))
paras_quantile<-as.data.frame(paras_quantile)
paras_quantile[1,4:8]<-c(0.020,0.2,0.5,1.2,1.6)
paras_quantile[2,4:8]<-c(0.05,0.4,1.6,3.0,3.8)
paras_quantile[3,4:8]<-c(0.05,0.4,6.5,15.0,20.0)
paras_quantile[4,4:8]<-c(0.01,0.05,0.1,0.3,0.5)
paras_quantile[5,4:8]<-c(0.02,0.1,0.4,0.6,0.9)
paras_quantile[6,4:8]<-c(0.05,0.4,3.3,4.9,6.0)
paras_quantile[7,4:8]<-c(0.005,0.02,0.05,0.2,0.3)
paras_quantile[8,4:8]<-c(0.01,0.02,0.1,0.2,0.4)
paras_quantile[9,4:8]<-c(0.02,0.2,0.7,1.3,1.6)

paras_quantile<-melt(paras_quantile,id.vars=c('Vs','sdMutation'),variable.name='quantile',value.name = 'speedCoef')

gradual_stats2_0<-gradual_stats2
gradual_stats2<-left_join(paras_quantile,gradual_stats2_0,by=c('Vs','sdMutation','speedCoef'))
```

### Lag between the new optimum and the population mean during the moving-optimum phase

```{r}
ggplot(gradual_stats2,aes(x=generation/N-10,y=gap))+
  geom_line(aes(group=speedCoef,col=speedCoef))+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Lag",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
gap_dy_p<-ggplot(gradual_stats2,aes(x=generation/N-10,y=gap/sdMutation))+
  geom_line(aes(group=speedCoef,col=speed/sdMutation/theta))+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_color_viridis_c(option="C")+
  # scale_x_continuous(trans='log10',labels=comma)+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Phenotypic\\ gap\\ (\\omega)"),col=TeX("$Speed\\ (\\omega\\theta)"))

gap_dy_p
```


### Phenotypic gap in extincted populations


```{r}
load("data/processed/stats2.RData")
extincted_stats2<-subset(extincted_stats2,speedCoef!=22)
extinct_dy_p<-ggplot(extincted_stats2,aes(x=generation/N-10,y=gap/sdMutation))+
  geom_line(aes(group=speedCoef,col=as.factor(round(speed/sdMutation/theta,2))),linewidth=1)+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_color_viridis_d(option = "C")+
  scale_x_continuous(limits=c(0,0.6))+
  scale_y_continuous(trans='log10')+ #,labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Phenotypic\\ gap\\ (\\omega)"),col=TeX("$Speed\\ (\\omega\\theta)"))

extinct_dy_p
```


### Fitness during the moving-optimum phase

```{r}
ggplot(gradual_stats2,aes(x=generation/N-10,y=meanFitness_mean))+
  geom_line(aes(group=speedCoef,col=speedCoef))+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Fitness",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
fit_dy_p<-ggplot(gradual_stats2,aes(x=generation/N-10,y=meanFitness_mean))+
  geom_line(aes(group=speedCoef,col=speed/sdMutation/theta))+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_color_viridis_c(option="C")+
  # scale_x_continuous(trans="log10",labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Fitness",col=TeX("$Speed\\ (\\omega\\theta)"))

fit_dy_p
```

### Genetic variance during the moving-optimum phase

```{r}
ggplot(gradual_stats2,aes(x=generation/N-10,y=varPheno_mean))+
  geom_line(aes(group=speedCoef,col=speedCoef))+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_color_viridis_c(option="C")+
  geom_hline(aes(yintercept=SSE_VG),alpha=.7,linetype='dashed')+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Genetic variance",
       col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
vg_dy_p<-ggplot(gradual_stats2,aes(x=generation/N-10,y=varPheno_mean/sdMutation^2))+
  geom_line(aes(group=speedCoef,col=speed/sdMutation/theta))+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_color_viridis_c(option="C")+
  geom_hline(aes(yintercept=SSE_VG/sdMutation^2),alpha=.7,linetype='dashed')+
  # scale_x_continuous(trans="log10",labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Genetic\\ variance\\ (\\omega^{2})"),
       col=TeX("$Speed\\ (\\omega\\theta)"))

vg_dy_p
```


# Dynamics of mutations during the moving-optimum phase



load data
```{r}
load("data/processed/mo_fixedMuta.RData")
```

### Fixation number ~ generation

```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=n_fixed,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Fixation number per generation",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
fn_dy_p<-ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=n_fixed,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  # scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Fixation number per generation",col=TeX("$Speed\\ (\\omega\\theta)"))

fn_dy_p
```

### Fixation time ~ generation
```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_fixedtime,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Fixation time",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_fixedtime,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Fixation time",col=TeX('$Speed\\ (\\omega\\theta)'))
```

### Mean fixed effect size ~ generation
```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_effect,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Mean fixed effect size",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
fe_dy_p<-ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_effect/sdMutation,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  # scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX('$Mean\\ fixed\\ effect\\ size\\ (\\omega)'),col=TeX('$Speed\\ (\\omega\\theta)'))

fe_dy_p
```


### Mean positive fixed effect ~ generation
```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_posEffect,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Mean positive fixed effect",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_posEffect/sdMutation,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX('$Mean\\ positive\\ fixed\\ effect\\ (\\omega)'),col=TeX('$Speed\\ (\\omega\\theta)'))
```


### Mean negative fixed effect ~ generation
```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_negEffect,group=speedCoef,col=speedCoef))+
  geom_point(size=.1)+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Mean negative fixed effect",col=TeX("$Speed\\ (\\sigma_{G(S)})"))

ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_negEffect,group=speedCoef,col=speedCoef))+
  geom_smooth()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Mean negative fixed effect",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_negEffect/sdMutation,group=speedCoef,col=speed/sdMutation/theta))+
  geom_point(size=.1)+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Mean\\ negative\\ fixed\\ effect (\\omega)"),col=TeX('$Speed\\ (\\omega\\theta)'))

ggplot(mo_fixedMuta3,aes(x=endTick_cut/N,y=mean_negEffect/sdMutation,group=speedCoef,col=speed/sdMutation/theta))+
  geom_smooth()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Mean\\ negative\\ fixed\\ effect\\ (\\omega)"),col=TeX('$Speed\\ (\\omega\\theta)'))
```


## Dynamics of segregating mutations during the moving-optimum phase

Calculate the mean of 100 replicates at the specific generation (record every 10 generations during the first and last 1000 generations and every 500 generations during the middle period)
```{r}
load("data/processed/muta2.RData")
shift_muta2_quantile<-left_join(paras_quantile,shift_muta2,by=c('Vs','sdMutation','speedCoef'))
mo_muta3_mean<-group_by(shift_muta2_quantile,Vs,sdMutation,speedCoef,generation)%>%filter(n()>=5)%>%
  summarise_at(vars(-replicate),mean,na.rm=T)
mo_muta3_mean<-left_join(mo_muta3_mean,parameters,by=c('Vs','sdMutation'))%>%mutate(speed=speedCoef*sqrt(SSE_VG))
```

### Segregating number ~ generation
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=n_segregations,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Segregating number per generation",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
sn_dy_p<-ggplot(mo_muta3_mean,aes(x=generation/N -10,y=n_segregations,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  # scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Segregating number per generation",col=TeX('$Speed\\ (\\omega\\theta)'))

sn_dy_p
```


### Proportion of positive segregating mutations ~ generation
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=n_posSegregations/n_segregations,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_y_continuous(labels=scales::percent)+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Proportion of positive fixations",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=n_posSegregations/n_segregations,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_y_continuous(labels=scales::percent)+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Proportion of positive fixations",col=TeX("$Speed\\ (\\omega\\theta)"))
```


### Mean segregating effect size ~ generation
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=mean_effect,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Mean segregating effect size",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
se_dy_p<-ggplot(mo_muta3_mean,aes(x=generation/N -10,y=mean_effect/sdMutation,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  # scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Mean\\ segregating\\ effect\\ size\\ (omega)"),col=TeX("$Speed\\ (\\omega\\theta)"))

se_dy_p
```

### Mean positive segregating effect ~ generation
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=mean_posEffect,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Mean positive segregating effect",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=mean_posEffect/sdMutation,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Mean\\ positive\\ segregating\\ effect\\ (omega)"),col=TeX("$Speed\\ (\\omega\\theta)"))
```

### Mean negative segregating effec size ~ generation
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=mean_negEffect,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Mean negative segregating effect",col=TeX("$Speed\\ (\\sigma_{G(S)})"))
```

Scaling by omega
```{r}
ggplot(mo_muta3_mean,aes(x=generation/N -10,y=mean_negEffect/sdMutation,group=speedCoef,col=speed/sdMutation/theta))+
  geom_line()+
  facet_grid(sdMutation~Vs,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)),scale='free_y')+
  scale_color_viridis_c(option="C")+
  scale_x_continuous(trans='log10',labels=comma)+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y=TeX("$Mean\\ negative\\ segregating\\ effect\\ (\\omega)"),col=TeX('$Speed\\ (\\omega\\theta)'))
```


### Supplementary figures
```{r}
pdf("dynamic_gap.pdf")
gap_dy_p+theme(aspect.ratio = 0.6)
dev.off()

pdf("dynamic_extinct_cases.pdf")
extinct_dy_p+theme(aspect.ratio = 0.6)
dev.off()

pdf("dynamic_fitness.pdf")
fit_dy_p+theme(aspect.ratio = 0.6)
dev.off()

pdf("dynamic_vg.pdf")
vg_dy_p+theme(aspect.ratio = 0.6)
dev.off()

pdf("dynamic_fixation_number.pdf")
fn_dy_p+theme(aspect.ratio = 0.6)
dev.off()

pdf("dynamic_fixed_effect.pdf")
fe_dy_p+theme(aspect.ratio = 0.6)
dev.off()

pdf("dynamic_segregating_number.pdf")
sn_dy_p+theme(aspect.ratio = 0.6)
dev.off()

pdf("dynamic_segregating_effect.pdf")
se_dy_p+theme(aspect.ratio = 0.6)
dev.off()
```


### Disconnect the database

```{r}
dbDisconnect(singleTrait_db)
```
