---
title: "Effect size distribution of mutations"
author: "Yuna"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: '2024-07-02'
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(echo = T,message =  F,warning=F,results = "hide")
```

### Working path

```{r}
setwd("/projects/ag-stetter/yzhang44/simulation/")
```

### Library packages

```{r}
library(DBI)
library(RSQLite)
library(dplyr)
library(reshape2)
library(ggplot2)
library(latex2exp)
library(ggpubr)# multiple figures in a page
library(scales)
library(cowplot)# get_legend() & plot_grid() functions
library(wesanderson)
library(stringr)
```

### Global plotting settings

```{r}
theme_set(theme_classic())
theme_update(text=element_text(size=13),plot.title=element_text(size=13),
             legend.text=element_text(size=10),legend.title=element_text(size=10,hjust=.5),
             legend.key.size=unit(.3,'cm'),
             legend.key.width=unit(.6,'cm'),
             legend.spacing=unit(0,'cm'),
             legend.background = element_rect(color="grey",fill=NA))

label_sdMutation<-function(string){TeX(paste("$\\omega : $",string))}

negativelog_trans <- function(base = exp(1)) {
    trans <- function(x) log(-x, base)
    inv <- function(x) -base^x
    trans_new(paste0("reverselog_", format(base)), transform = trans, inverse=inv)
}
```


## Effect size distribution of fixations

### case from trait of Vs=1 & sigma_m=0.05
```{r}
N=10000
Vs=1
sd=0.05

sse_fixedMuta100<-data.frame()
for(replicate in 1:100){
  fixedMuta<-read.table(paste0("Vs",Vs,"sd",sd,"/burn_in/fixedMuta_",replicate,".txt"),
                        colClasses=c("integer","NULL","NULL","numeric","integer","integer"),
                        col.names=c("replicate","id","position","effectSize","aroseTick","endTick"))
  sse_fixedMuta100<-rbind(sse_fixedMuta100,fixedMuta)
}

speedCoef=0.5
moe_fixedMuta100<-data.frame()
for(replicate in 1:100){
  fixedMuta<-read.table(paste0("Vs",Vs,"sd",sd,"/gradual/fixedMuta_",speedCoef,"_",replicate,".txt"),
                        colClasses=c("integer","NULL","NULL","numeric","integer","integer"),
                        col.names=c("replicate","id","position","effectSize","aroseTick","endTick"))
  moe_fixedMuta100<-rbind(moe_fixedMuta100,fixedMuta)
}
moe_fixedMuta100$period<-ifelse(moe_fixedMuta100$endTick>10.1*N,"0.1-10N",
                                ifelse(moe_fixedMuta100$endTick>10.05*N,"0.05-0.1N",
                                       ifelse(moe_fixedMuta100$endTick>10.02*N,"0.02-0.05",
                                              ifelse(moe_fixedMuta100$endTick>10.01*N,"0.01-0.02N","0-0.01N"))))

# overall distribution
plot(density(moe_fixedMuta100$effectSize),col='red',ylim=c(0,30),main='Distribution of effect size',xlab='Effect size',lwd=2)
lines(density(sse_fixedMuta100$effectSize),lty=2,lwd=2)
lines(seq(-1,1,length=1000),dnorm(seq(-1,1,length=1000),0,0.05),lty=3,lwd=2)
legend(0.15,30,legend=c("Moving optimum","Fixed optimum","Supply"),lty=1:3,col=c('red','black','black'),lwd=2)

# different periods
ggplot()+
  geom_density(data=moe_fixedMuta100,aes(x=effectSize,col=period))+
  # facet_grid(period~.)+
  geom_density(data=sse_fixedMuta100,aes(x=effectSize),col='black',linetype='dashed')+
  geom_line(data=data.frame(x=seq(-0.1,0.3,length=500),y=dnorm(seq(-0.1,0.3,length=500),0,0.05)),aes(x,y),linetype='dotted')+
  labs(x="Effect size",y="Density",col='Period')

# different periods with fixations arisen before shift or not
ggplot()+
  geom_density(data=moe_fixedMuta100,aes(x=effectSize,fill=aroseTick>10*N,col=aroseTick>10*N),alpha=.2)+
  facet_grid(period~.)+
  geom_density(data=sse_fixedMuta100,aes(x=effectSize),col='black',linetype='dashed')+
  geom_line(data=data.frame(x=seq(-0.1,0.3,length=500),y=dnorm(seq(-0.1,0.3,length=500),0,0.05)),aes(x,y),linetype='dotted')+
  labs(x="Effect size",y="Density",fill="After shift",col='After shift')
```

### effect size distribution of all fixations
```{r}
singleTrait_db<-dbConnect(RSQLite::SQLite(),dbname="~/shortcut/B6/simulation/singleTrait.sqlite")
on.exit(dbDisconnect(singleTrait_db))

dbListTables(singleTrait_db)

parameters<-tbl(singleTrait_db,"parameters")%>%as.data.frame
sse_fixedmuta<-tbl(singleTrait_db,"burnin_fixedMuta")%>%as.data.frame  #15016
moe_fixedMuta<-tbl(singleTrait_db,"gradual_fixedMuta")
# moe_fixedMuta<-as.data.frame(moe_fixedMuta)
moe_fixedMuta1<-filter(moe_fixedMuta,replicate==1)%>%as.data.frame  # 2073928
moe_fixedMuta1<-left_join(moe_fixedMuta1,parameters,by=c("Vs","sdMutation"))
moe_fixedMuta1<-mutate(moe_fixedMuta1,speed=speedCoef*sqrt(SSE_VG))


summ<-group_by(moe_fixedMuta1,Vs,sdMutation,speed/sdMutation)%>%
  summarise(n_fixed=n())

supply_def<-rbind(data.frame(effectSize=seq(-2,2,length=500),density=dnorm(seq(-2,2,length=500),mean=0,sd=0.05),sdMutation=0.05),
                  data.frame(effectSize=seq(-2,2,length=500),density=dnorm(seq(-2,2,length=500),mean=0,sd=0.1),sdMutation=0.1),
                  data.frame(effectSize=seq(-2,2,length=500),density=dnorm(seq(-2,2,length=500),mean=0,sd=0.5),sdMutation=0.5))
supply_def<-rbind(data.frame(supply_def,Vs=1),
                  data.frame(supply_def,Vs=10),
                  data.frame(supply_def,Vs=100))

ggplot()+
  geom_area(data=supply_def,aes(x=effectSize,y=density),fill='snow3')+
  geom_density(data=sse_fixedmuta,aes(x=effectSize))+
  facet_grid(Vs~sdMutation,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density")

pal <- wes_palette("Zissou1", 100, type = "continuous")

ggplot()+
  geom_density(data=moe_fixedMuta1,aes(x=effectSize,group=speedCoef,col=speed/sdMutation),trim=T)+
  scale_color_gradientn(colours = pal)+
  facet_grid(Vs~sdMutation,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density",col=TeX("$Speed\\ (\\sigma_m)"))

ggplot()+
  geom_area(data=supply_def,aes(x=effectSize,y=density),fill='snow3')+
  geom_density(data=sse_fixedmuta,aes(x=effectSize))+
  geom_density(data=moe_fixedMuta1,aes(x=effectSize,group=speedCoef,col=speed/sdMutation),trim=T)+
  scale_color_gradientn(colours = pal)+
  facet_wrap(Vs~sdMutation,scales="free",labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density",col=TeX("$Speed\\ (\\sigma_m)"))


ggplot()+
  geom_density(data=moe_fixedMuta1,aes(x=effectSize/sdMutation,col=factor(round(speed/sdMutation,2))))+
  geom_density(data=sse_fixedmuta,aes(x=effectSize/sdMutation),linetype='dashed')+
  geom_line(data=supply_def,aes(x=effectSize/sdMutation,y=density*sdMutation),linetype='dotted')+
  facet_grid(Vs~sdMutation)

ggplot(supply_def,aes(x=effectSize/sdMutation,y=density*sdMutation))+
  geom_line()+
  facet_wrap(Vs~sdMutation,scales='free')



moe_fixedMuta_Vs100sd0.05<-filter(moe_fixedMuta,Vs==100 & sdMutation==0.05)%>%as.data.frame
moe_fixedMuta_Vs100sd0.05<-left_join(moe_fixedMuta_Vs100sd0.05,parameters,by=c("Vs","sdMutation"))
moe_fixedMuta_Vs100sd0.05<-mutate(moe_fixedMuta_Vs100sd0.05,speed=speedCoef*sqrt(SSE_VG))
ggplot(moe_fixedMuta_Vs100sd0.05,aes(x=effectSize,col=factor(round(speed/sdMutation,2))))+
  geom_density()

moe_fixedMuta_Vs1sd0.5<-filter(moe_fixedMuta,Vs==1 & sdMutation==0.5)%>%as.data.frame
moe_fixedMuta_Vs1sd0.5<-left_join(moe_fixedMuta_Vs1sd0.5,parameters,by=c("Vs","sdMutation"))
moe_fixedMuta_Vs1sd0.5<-mutate(moe_fixedMuta_Vs1sd0.5,speed=speedCoef*sqrt(SSE_VG))
ggplot(moe_fixedMuta_Vs1sd0.5,aes(x=effectSize,col=factor(round(speed/sdMutation,2))))+
  geom_density()
```



## Effect size distribution of segregating mutations

### Supply distribution vs. stabilizing selection
```{r}
singleTrait_db<-dbConnect(RSQLite::SQLite(),dbname="~/shortcut/B6/simulation/singleTrait.sqlite")
on.exit(dbDisconnect(singleTrait_db))

dbListTables(singleTrait_db)

N=10000

parameters<-tbl(singleTrait_db,"parameters")%>%as.data.frame

sse_muta<-tbl(singleTrait_db,"burnin_muta")%>%
  filter(generation==10*N)%>%as.data.frame     # 189432

supply_def<-rbind(data.frame(effectSize=seq(-2,2,length=500),density=dnorm(seq(-2,2,length=500),mean=0,sd=0.05),sdMutation=0.05),
                  data.frame(effectSize=seq(-2,2,length=500),density=dnorm(seq(-2,2,length=500),mean=0,sd=0.1),sdMutation=0.1),
                  data.frame(effectSize=seq(-2,2,length=500),density=dnorm(seq(-2,2,length=500),mean=0,sd=0.5),sdMutation=0.5))
supply_def<-rbind(data.frame(supply_def,Vs=1),
                  data.frame(supply_def,Vs=10),
                  data.frame(supply_def,Vs=100))

# without weights of allelic frequency
ggplot()+
    geom_area(data=supply_def,aes(x=effectSize,y=density),fill='snow3')+
    geom_density(data=sse_muta,aes(x=effectSize))+
    facet_grid(Vs~sdMutation,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
    scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density")


# with weights of allelic frequency
sse_weighted_density<-split(sse_muta,paste0(sse_muta$Vs,"_",sse_muta$sdMutation))%>%
  lapply(FUN=function(x){density(x$effectSize,weights=x$frequency/sum(x$frequency))})

sse_muta_density<-data.frame()
for(i in 1:length(sse_weighted_density)){
  Vs = word(names(sse_weighted_density[i]),1,sep="_")
  sdMutation = word(names(sse_weighted_density[i]),2,sep="_")
  
  df<-data.frame(Vs=Vs,sdMutation=sdMutation,effectSize=sse_weighted_density[[i]]$x,density=sse_weighted_density[[i]]$y)
  sse_muta_density<-rbind(sse_muta_density,df)
}
sse_muta_density<-mutate_at(sse_muta_density,.vars = c("Vs","sdMutation"),.funs = as.numeric)

ggplot()+
  geom_area(data=supply_def,aes(x=effectSize,y=density),fill='snow3')+
  geom_line(data=sse_muta_density,aes(x=effectSize,y=density))+
  facet_grid(Vs~sdMutation,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  # scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density")
```

```{r}
moe_muta<-tbl(singleTrait_db,"gradual_muta")%>%
  filter(generation==20*N)%>%as.data.frame

 # moe_muta<-left_join(moe_muta,parameters,by=c("Vs","sdMutation"))
# moe_muta<-mutate(moe_muta,speed=speedCoef*sqrt(SSE_VG))

pal <- wes_palette("Zissou1", 100, type = "continuous")

ggplot()+
  geom_density(data=moe_muta,aes(x=effectSize,group=speedCoef,col=speed/sdMutation),trim=T)+
  scale_color_gradientn(colours = pal)+
  facet_grid(Vs~sdMutation,labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density",col=TeX("$Speed\\ (\\sigma_m)"))

ggplot()+
  geom_area(data=supply_def,aes(x=effectSize,y=density),fill='snow3')+
  geom_density(data=sse_muta,aes(x=effectSize))+
  geom_density(data=moe_muta,aes(x=effectSize,group=speedCoef,col=speed/sdMutation),trim=T)+
  scale_color_gradientn(colours = pal)+
  facet_wrap(Vs~sdMutation,scales="free",labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density",col=TeX("$Speed\\ (\\sigma_m)"))


# with weights of allelic frequency
moe_weighted_density<-split(moe_muta,paste0(moe_muta$Vs,"_",moe_muta$sdMutation,"_",moe_muta$speedCoef))%>%
  lapply(FUN=function(x){density(x$effectSize,weights=x$frequency/sum(x$frequency))})

moe_muta_density<-data.frame()
for(i in 1:length(moe_weighted_density)){
  Vs = word(names(moe_weighted_density[i]),1,sep="_")
  sdMutation = word(names(moe_weighted_density[i]),2,sep="_")
  speedCoef = word(names(moe_weighted_density[i]),3,sep="_")
  
  df<-data.frame(Vs=Vs,sdMutation=sdMutation,speedCoef=speedCoef,effectSize=moe_weighted_density[[i]]$x,density=moe_weighted_density[[i]]$y)
  moe_muta_density<-rbind(moe_muta_density,df)
}
moe_muta_density<-mutate_at(moe_muta_density,.vars = c("Vs","sdMutation","speedCoef"),.funs = as.numeric)

moe_muta_density<-left_join(moe_muta_density,parameters,by=c("Vs","sdMutation"))
moe_muta_density<-mutate(moe_muta_density,speed=speedCoef*sqrt(SSE_VG))

ggplot()+
  geom_area(data=supply_def,aes(x=effectSize,y=density),fill='snow3')+
  geom_line(data=sse_muta_density,aes(x=effectSize,y=density))+
  geom_line(data=moe_muta_density,aes(x=effectSize,y=density,group=speedCoef,col=speed/sdMutation))+
  scale_color_gradientn(colours = pal)+
  facet_wrap(Vs~sdMutation,scales="free",labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  # scale_x_continuous(limits=c(-0.5,1.5))+
  labs(x="Effect size",y="Density",col=TeX("$Speed\\ (\\sigma_m)"))
```


### Mean and sd. of the effect sizes of segregating site (with allelic frequencies)
```{r}
sse_muta_weightedMean<-split(sse_muta,paste0(sse_muta$Vs,"_",sse_muta$sdMutation))%>%
  lapply(FUN=function(x){
    mean_effect<-sum(x$frequency * x$effectSize) / sum(x$frequency)
    sd_effect<-sqrt(sum(x$frequency * (x$effectSize - mean_effect)^2) / sum(x$frequency))
    return(c(mean_effect,sd_effect))
  })

sse_weightedMeanSd<-data.frame()
for(i in 1:length(sse_muta_weightedMean)){
  Vs = word(names(sse_muta_weightedMean[i]),1,sep="_")
  sdMutation = word(names(sse_muta_weightedMean[i]),2,sep="_")
  
  df<-data.frame(Vs=Vs,sdMutation=sdMutation,mean_effect=sse_muta_weightedMean[[i]][1],sd_effect=sse_muta_weightedMean[[i]][2])
  sse_weightedMeanSd<-rbind(sse_weightedMeanSd,df)
}
sse_weightedMeanSd<-mutate_at(sse_weightedMeanSd,.vars = c("Vs","sdMutation"),.funs = as.numeric)

###
paras
moe_muta_weightedMean<-tbl(singleTrait_db,"gradual_muta")%>%
  filter(generation>19.9*N)%>%
  group_by(Vs,sdMutation,speedCoef)%>%
  summarise(mean_effect=sum(frequency * effectSize)/ sum(frequency))
moe_muta_weightedMean<-as.data.frame(moe_muta_weightedMean)

moe_muta_weightedMean<-split(moe_muta,paste0(moe_muta$Vs,"_",moe_muta$sdMutation,"_",moe_muta$speedCoef))%>%
  lapply(FUN=function(x){
    mean_effect<-sum(x$frequency * x$effectSize) / sum(x$frequency)
    sd_effect<-sqrt(sum(x$frequency * (x$effectSize - mean_effect)^2) / sum(x$frequency))
    return(c(mean_effect,sd_effect))
  })

moe_weightedMeanSd<-data.frame()
for(i in 1:length(moe_muta_weightedMean)){
  Vs = word(names(moe_muta_weightedMean[i]),1,sep="_")
  sdMutation = word(names(moe_muta_weightedMean[i]),2,sep="_")
  speedCoef = word(names(moe_muta_weightedMean[i]),3,sep="_")
  
  df<-data.frame(Vs=Vs,sdMutation=sdMutation,speedCoef=speedCoef,mean_effect=moe_muta_weightedMean[[i]][1],sd_effect=moe_muta_weightedMean[[i]][2])
  moe_weightedMeanSd<-rbind(moe_weightedMeanSd,df)
}
moe_weightedMeanSd<-mutate_at(moe_weightedMeanSd,.vars = c("Vs","sdMutation","speedCoef"),.funs = as.numeric)
moe_weightedMeanSd<-left_join(moe_weightedMeanSd,parameters,by=c("Vs","sdMutation"))%>%
  mutate(speed=speedCoef * sqrt(SSE_VG),
         sigmaOmega2=sdMutation^2/2/Vs,
         group=paste0("Vs=",Vs,", $\\sigma_{m}=$",sdMutation))

so=unique(moe_weightedMeanSd$sigmaOmega2)%>%sort
so=as.character(format(so,scientific = T))
so[5:6]<-c("1.25e-03_a","1.25e-03_b")

moe_weightedMeanSd$sigmaOmega2<-format(moe_weightedMeanSd$sigmaOmega2,scientific=T)

moe_weightedMeanSd[which(moe_weightedMeanSd$Vs==1 & moe_weightedMeanSd$sdMutation==0.05),"sigmaOmega2"]<-"1.25e-03_a"
moe_weightedMeanSd[which(moe_weightedMeanSd$Vs==100 & moe_weightedMeanSd$sdMutation==0.5),"sigmaOmega2"]<-"1.25e-03_b"
moe_weightedMeanSd$sigmaOmega2<-factor(moe_weightedMeanSd$sigmaOmega2,level=so,label=so)

save(sse_weightedMeanSd,moe_weightedMeanSd,file="muta_weightedMeanSd.RData")
```

### Weighted mean vs. nonweighted mean (last 1000 generations) vs. nonweighted mean at 20N generation
```{r}
load("muta2(fill0).RData")
load("muta_weightedMeanSd.RData")
moe_weightedMeanSd_20N<-moe_weightedMeanSd
sse_weightedMeanSd_20N<-sse_weightedMeanSd
rm(moe_weightedMeanSd,sse_weightedMeanSd)
load("summ_weighted.RData")

N=10000 

# mean at 20 N generation
moe_muta_mean20N<-subset(shift_muta2,generation==20*N)%>%
  group_by(Vs,sdMutation,speedCoef,replicate)%>%
  summarise_at(vars(-generation),mean,na.rm=T)%>%
  group_by(Vs,sdMutation,speedCoef)%>%summarise_at(vars(-replicate),mean,na.rm=T)

# mean during 19.9 ~ 20 N generation
moe_muta_mean<-subset(shift_muta2,generation>19.9*N)%>%
  group_by(Vs,sdMutation,speedCoef,replicate)%>%
  summarise_at(vars(-generation),mean,na.rm=T)%>%
  group_by(Vs,sdMutation,speedCoef)%>%summarise_at(vars(-replicate),mean,na.rm=T)

colnames(moe_muta_mean20N)[5]<-"mean_effect20N"
colnames(moe_weightedMeanSd_20N)[4]<-"weighted_mean_effect"

joint_df<-left_join(moe_weightedMeanSd_20N[,c("Vs","sdMutation","speedCoef","weighted_mean_effect","sd_effect")],
                    moe_muta_mean[,c("Vs","sdMutation","speedCoef","mean_effect")],by=c("Vs","sdMutation","speedCoef"))%>%left_join(moe_muta_mean20N[,c("Vs","sdMutation","speedCoef","mean_effect20N")],by=c("Vs","sdMutation","speedCoef"))
joint_df<-left_join(joint_df,parameters,by=c("Vs","sdMutation"))%>%
  mutate(speed=speedCoef * sqrt(SSE_VG),
         sigmaOmega2=sdMutation^2/2/Vs,
         group=paste0("Vs=",Vs,", $\\sigma_{m}=$",sdMutation))

ggplot(joint_df,aes(x=mean_effect20N,y=weighted_mean_effect,group=group,col=factor(sigmaOmega2)))+
  geom_point()+
  geom_abline(slope=1,intercept=0)+
  labs(x="at 20N",y="weighted")
ggplot(joint_df,aes(x=mean_effect,y=weighted_mean_effect,group=group,col=factor(sigmaOmega2)))+
  geom_point()+
  geom_abline(slope=1,intercept=0)+
  labs(x="last 1000 generations",y="weighted")
ggplot(joint_df,aes(x=mean_effect,y=mean_effect20N,group=group,col=factor(sigmaOmega2)))+
  geom_point()+
  geom_abline(slope=1,intercept=0)+
  labs(x="last 1000 generations",y="at 20N")
```

### Weighted mean of the last 1000 generations vs. mean of the weighted mean of the last 1000 generations 
```{r}
head(moe_muta_weightedMeanSd_generation)
ggplot(moe_muta_weightedMeanSd_generation,aes(x=generation,y=mean_effect,group=speedCoef,col=speedCoef))+
  geom_line()+
  facet_wrap(Vs~sdMutation,scales='free')

# group_by(moe_muta_weightedMeanSd_generation,Vs,sdMutation,speedCoef)%>%
#   summarise(var_amongGeneration=var(mean_effect))%>%ungroup%>%select(var_amongGeneration)%>%unlist

# mean of the summary statistics of the last 1000 generations
mean_effect_amongGeneration<-group_by(moe_muta_weightedMeanSd_generation,Vs,sdMutation,speedCoef)%>%
  summarise(mean_effect=mean(mean_effect),sd_effect=mean(sd_effect))

weighted1000_vs_mean1000<-left_join(mean_effect_amongGeneration,moe_muta_weightedMeanSd,by=c("Vs","sdMutation","speedCoef"))

plot(weighted1000_vs_mean1000$mean_effect.x,weighted1000_vs_mean1000$mean_effect.y)
abline(0,1)
plot(weighted1000_vs_mean1000$sd_effect.x,weighted1000_vs_mean1000$sd_effect.y)
abline(0,1)
```
They are very closed. So,use weighted mean of the last 1000 generations

### weighted mean of the last 1000 generations vs. weighted mean of the last generation
```{r}
last1000_vs_last<-left_join(moe_weightedMeanSd_20N,moe_muta_weightedMeanSd,by=c('Vs','sdMutation','speedCoef'))

plot(sqrt(last1000_vs_last$sd_effect.x) , last1000_vs_last$sd_effect.y)
abline(0,1)
```
Few points are a bit off in the high sd_effect region. So, use the weighted mean of the last 1000 generations

### weighted mean vs. unweighted mean of the last 1000 generations
```{r}
weighted_vs_unweighted<-left_join(moe_muta_mean,moe_muta_weightedMeanSd,by=c('Vs','sdMutation','speedCoef'))

plot(weighted_vs_unweighted$mean_effect.x,weighted_vs_unweighted$mean_effect.y)
abline(0,1)
```
Weighted mean is always larger than the unweighted mean

number of segregating sites
```{r}
moe_muta_weightedMeanSd_generation$n_segregations<-moe_muta_weightedMeanSd_generation$n_segregations / 100
num<-group_by(moe_muta_weightedMeanSd_generation,Vs,sdMutation,speedCoef)%>%
  summarise(n_segregations=sum(n_segregations)/n())
num_vs<-left_join(num,moe_muta_mean,by=c('Vs','sdMutation','speedCoef'))

plot(num_vs$n_segregations.x,num_vs$n_segregations.y)
abline(0,1)
```
Those speeds with replicates less than 100 are not accurate in the weighted data frame.

### mean effect of fixations vs. segregating sites
```{r}
fix_vs_segre<-left_join(moe_fixed2,moe_muta_weightedMeanSd,by=c('Vs','sdMutation','speedCoef'))

plot(fix_vs_segre$mean_effect.x,fix_vs_segre$mean_effect.y,xlab="mean effect of fixations",ylab="mean effect of segregating")
abline(0,1)

plot(fix_vs_segre$sd_effect.x,fix_vs_segre$sd_effect.y,xlab="std. of effect of fixations",ylab="std. of effect of segregating")
abline(0,1)

sd_diff<-mutate(fix_vs_segre,sd_effect_diff=sd_effect.x - sd_effect.y)
sd_diff[order(sd_diff$sd_effect_diff,decreasing = T),]
ggplot(sd_diff,aes(x=speed/sdMutation,y=sd_effect_diff/sdMutation,group=group,col=sigmaOmega2))+
  scale_color_manual(name=TeX("$Trait\\ (\\frac{omega^{2}}{2V_s})"),values=legend_pattern$color,labels=legend_pattern$sigmaOmega2)+
  geom_point()+
  geom_line()+
  geom_hline(yintercept=0)+
  labs(x=TeX("$Speed\\ (\\sigma)"),y=TeX("$Sd(fixations) - sd(segregating)\\ (\\sigma)"))
```
Effect size of them are highly related, but the effect size of fixations are larger.
Std. of effect size are also highly related. In general, segregating sites have higher std., but, in the second largest part, fixations have larger std.



### Allelic frequency ~ effect size matrix
```{r eval=F}
N=10000

singleTrait_db<-dbConnect(RSQLite::SQLite(),dbname="/projects/ag-stetter/yzhang44/simulation/singleTrait.sqlite")
on.exit(dbDisconnect(singleTrait_db))

dbListTables(singleTrait_db)

N=10000

parameters<-tbl(singleTrait_db,"parameters")%>%as.data.frame

segre_stable<-tbl(singleTrait_db,"burnin_muta")%>%
  filter(Vs==1 & sdMutation==0.5 & generation==10*N)%>%
  as.data.frame

segre_case<-tbl(singleTrait_db,"gradual_muta")%>%
  filter(Vs==1 & sdMutation==0.5 & speedCoef %in% c(0.002,0.5,1.0,6.5,10.0,13.5,18.0,21.0) & generation==20*N)%>%
  as.data.frame
# 0.5 is where the max segregating number & the mean effect close to sdMutation=0.5
# 1.0 is where the speed is 1 sqrt(SSE_VG)
# 10.0 is where the speed is 1 sdMutation=0.5 & max weighted mean_effect
# 6.5 is where the max sd_effect, max genetic variance
# 18.0 is where the sd_effect is at the bottom of fast speeds
# 0.002 & 21.0 are the slowest and fastest speed (less than 100 replicates at speedCoef=21.5)

segre_case<-rbind(mutate(segre_stable,speedCoef=0),segre_case)
# quantile(segre_case$effectSize)
#         0%        25%        50%        75%       100% 
# -1.7177300 -0.0108042  0.2692030  0.5657222  2.0909700 
# quantile(segre_case$frequency)
#      0%     25%     50%     75%    100% 
# 0.00005 0.00005 0.00050 0.02525 0.99995 

segre_case<-mutate(segre_case,Vs=1,sdMutation=0.5)
save(segre_case,file="segre_case.RData")
```

add stabilizing state as a contrast
```{r}
load("data/processed/segre_case.RData")
segre_case<-mutate(segre_case,effectSize_sdMutation=effectSize/sdMutation)
segre_case$freq_left<-cut(segre_case$frequency,breaks=c(0,5e-05,1e-04,1e-03,1e-02,seq(0.1,1,0.1)),labels=F)
segre_case$eff_left<-cut(segre_case$effectSize_sdMutation,breaks=seq(-4,4.2,0.2),labels=F)

summ_segre_case<-group_by(segre_case,speedCoef)%>%
  summarise(mean_effect=sum(frequency * effectSize)/sum(frequency),
            sd_effect=sqrt(sum(frequency * (effectSize - mean_effect)^2) / sum(frequency)))

num_df<-group_by(segre_case,speedCoef)%>%
  summarise(total_num=n(),n_replicate=length(unique(replicate)))
num_df<-mutate(num_df,n_segre=round(total_num/n_replicate,2))
segre_case_matrix<-group_by(segre_case,speedCoef,eff_left,freq_left)%>%summarise(num=n())%>%
  left_join(.,num_df,by=c('speedCoef'))%>%
  mutate(relFreq=num/total_num)

subdf<-subset(segre_case_matrix,speedCoef %in% c(0,0.5,10))
subdf$speedCoef<-factor(subdf$speedCoef,levels=c(0,0.5,10),labels=c("Before shift","Slow shift","Fast shift"))
sub_num_df<-subset(num_df,speedCoef %in% c(0,0.5,10))
sub_num_df$speedCoef<-factor(sub_num_df$speedCoef,levels=c(0,0.5,10),labels=c("Before shift","Slow shift","Fast shift"))
matrix_p<-ggplot()+
  geom_rect(data=subdf,aes(xmin=eff_left,xmax=eff_left+1,ymin=freq_left,ymax=freq_left+1,fill=relFreq))+
  scale_fill_viridis_c(trans='log10',name="Proportion",breaks=c(1e-4,1e-3,1e-2),labels=TeX(c("$10^{-4}","$10^{-3}","$10^{-2}")))+
  facet_grid(speedCoef~.)+
  geom_vline(xintercept = 21,col='red')+
  scale_x_continuous(breaks=c(1,11,21,31,41),labels=c(-4,-2,0,2,4))+
  scale_y_continuous(breaks=c(1,2,3,4,5,6,10,14),labels=TeX(c("0","$5\\times 10^{-5}","$10^{-4}","$10^{-3}","$10^{-2}","0.1","0.5","0.9")))+
  geom_text(data=sub_num_df,aes(x=9,y=12,label=paste0("n=",n_segre)))+
  labs(x=TeX("$Effect\\ size\\ (\\omega)"),y="Allele frequency")+
  theme(panel.grid.major = element_line(),legend.position="bottom",aspect.ratio = 0.5)

pdf("output/matrix_polygenic.pdf",width=5,height=6)
matrix_p
dev.off()

## remove new mutations
segre_case_nonNew<-subset(segre_case,aroseTick!=20*N)
segre_case_nonNew<-subset(segre_case_nonNew, !(speedCoef==0 & aroseTick==10*N))

num_df_nonNew<-group_by(segre_case_nonNew,speedCoef)%>%summarise(total_num=n())
segre_case_matrix_nonNew<-group_by(segre_case_nonNew,speedCoef,eff_left,freq_left)%>%summarise(num=n())%>%
  left_join(.,num_df_nonNew,by=c('speedCoef'))%>%
  mutate(relFreq=num/total_num)

ggplot(subset(segre_case_matrix_nonNew,speedCoef %in% c(0,0.5,10)))+
  geom_rect(aes(xmin=eff_left,xmax=eff_left+1,ymin=freq_left,ymax=freq_left+1,fill=relFreq))+
  scale_fill_viridis_c(trans="log10")+
  facet_grid(speedCoef~.)+
  geom_vline(xintercept = 21,col='red')+
  scale_x_continuous(breaks=c(1,11,21,31,41),labels=c(-4,-2,0,2,4))+
  scale_y_continuous(breaks=c(1,2,3,4,5,6,10,14),labels=TeX(c("0","$5\\times 10^{-5}","$10^{-4}","$10^{-3}","$10^{-2}","0.1","0.5","0.9")))+
  labs(x=TeX("$Effect\\ size\\ (\\omega)"),y="Allele frequency")+
  theme(panel.grid.major = element_line(),legend.position="bottom",aspect.ratio = 0.5)
```



### Allelic frequency ~ effect size matrix of all traits
```{r eval=F}
# data are from summary_statistics.Rmd
maxVG<-split(df,df$group)%>%lapply(.,FUN=function(x){return(x[which.max(x$MOE_VG),])})%>%do.call('rbind',.)%>%as.data.frame; rownames(maxVG)<-NULL

maxSN<-split(moe_muta2,moe_muta2$group)%>%lapply(.,FUN=function(x){return(x[which.max(x$n_segregations),])})%>%do.call('rbind',.)%>%as.data.frame; rownames(maxSN)<-NULL

maxSE<-split(moe_muta2,moe_muta2$group)%>%lapply(.,FUN=function(x){return(x[which.max(x$mean_effect),])})%>%do.call('rbind',.)%>%as.data.frame; rownames(maxSE)<-NULL

maxFE<-split(moe_fixed2,moe_fixed2$group)%>%lapply(.,FUN=function(x){return(x[which.max(x$mean_effect),])})%>%do.call('rbind',.)%>%as.data.frame; rownames(maxFE)<-NULL

# maxSEW<-split(moe_muta_weightedMeanSd,moe_muta_weightedMeanSd$group)%>%lapply(.,FUN=function(x){return(x[which.max(x$mean_effect),])})%>%do.call('rbind',.)%>%as.data.frame; rownames(maxSEW)<-NULL 
# from the old simulations: only Vs1sd0.5 & Vs10sd0.1 are close, others are the same)

maxSPEED<-group_by(paras,Vs,sdMutation)%>%slice_max(order_by = speedCoef,n=2)
minSPEED<-group_by(paras,Vs,sdMutation)%>%slice_min(order_by = speedCoef)

paras_vg<-left_join(paras,parameters,by=c("Vs","sdMutation"))
paras_vg<-mutate(paras_vg,speed_omega_1=abs(speedCoef*sqrt(SSE_VG)/sdMutation-1))
OneOmegaSPEED<-group_by(paras_vg,Vs,sdMutation)%>%slice_min(order_by = speed_omega_1)
OneOmegaSPEED<-OneOmegaSPEED[,c("Vs","sdMutation","speedCoef")]
OneSigmaVGSPPED<-paras[paras$speedCoef==0.5,]

matrix_speeds<-rbind(maxVG[,1:3],maxSN[,1:3],maxSE[,1:3],maxFE[,1:3],maxSPEED[,1:3],minSPEED[,1:3],OneOmegaSPEED,OneSigmaVGSPPED)
matrix_speeds<-matrix_speeds[order(matrix_speeds$Vs,matrix_speeds$sdMutation,matrix_speeds$speedCoef),]

matrix_speeds<-unique(matrix_speeds)# 59 * 3
save(matrix_speeds,file="matrix_speeds.RData")
```

```{r eval=F}
singleTrait_db<-dbConnect(RSQLite::SQLite(),dbname="~/shortcut/B6/simulation/singleTrait.sqlite")
on.exit(dbDisconnect(singleTrait_db))

dbListTables(singleTrait_db)

N=10000

parameters<-tbl(singleTrait_db,"parameters")%>%as.data.frame

segre_stable<-tbl(singleTrait_db,"burnin_muta")%>%
  filter(generation==10*N)%>%
  as.data.frame

segre_speeds<-c()
for(Vs0 in c(1,10,100)){
  for(sdMutation0 in c(0.05,0.1,0.5)){
    speedCoefs<-subset(matrix_speeds,Vs==Vs0 & sdMutation==sdMutation0)$speedCoef
    
    segre_case<-tbl(singleTrait_db,"gradual_muta")%>%
      filter(Vs==Vs0 & sdMutation==sdMutation0 & speedCoef %in% speedCoefs & generation==20*N)%>%
      as.data.frame
    
    segre_speeds<-rbind(segre_speeds,segre_case)
  }
}

segre<-rbind(mutate(segre_stable,speedCoef=0),segre_speeds)
save(parameters,segre,file="segre.RData")
```

```{r}
load("data/processed/segre.RData")

segre<-left_join(segre,parameters,by=c("Vs","sdMutation"))
segre<-mutate(segre,effectSize_sdMutation=effectSize/sdMutation,
              speed_omega=round(speedCoef*sqrt(SSE_VG)/sdMutation,4))
segre$freq_left<-cut(segre$frequency,breaks=c(0,1e-04,1e-03,1e-02,seq(0.1,1,0.1)),label=F)
segre$eff_left<-cut(segre$effectSize_sdMutation,breaks=seq(-4.4,4.4,0.2),labels=F)

num_df<-group_by(segre,Vs,sdMutation,speedCoef,speed_omega)%>%
  summarise(total_num=n(),n_replicate=length(unique(replicate)))
num_df<-mutate(num_df,n_segre=round(total_num/n_replicate,2))
segre_matrix<-group_by(segre,Vs,sdMutation,speedCoef,speed_omega,freq_left,eff_left)%>%summarise(num=n())%>%
  left_join(.,num_df,by=c('Vs','sdMutation','speedCoef','speed_omega'))%>%
  mutate(relFreq=num/total_num)


label_speed<-function(string){TeX(paste(string,"$\\omega$"))}

for(Vs0 in c(1,10,100)){
  for(sdMutation0 in c(0.05,0.1,0.5)){
    subdf<-subset(segre_matrix,Vs==Vs0 & sdMutation==sdMutation0)
    sub_num_df<-subset(num_df,Vs==Vs0 & sdMutation==sdMutation0)
    
    p_title=paste0("Trait: Vs=",Vs0," & $\\omega=",sdMutation0)
    
    matrix_p<-ggplot()+
      geom_rect(data=subdf,aes(xmin=eff_left,xmax=eff_left+1,ymin=freq_left,ymax=freq_left+1,fill=relFreq))+
      scale_fill_viridis_c(trans='log10',name="Proportion",breaks=c(1e-4,1e-3,1e-2),
                           labels=TeX(c("$10^{-4}","$10^{-3}","$10^{-2}")))+
      facet_grid(speed_omega~.,labeller=labeller(speed_omega=as_labeller(label_speed,label_parsed)))+
      geom_vline(xintercept = 23,col='red')+
      scale_x_continuous(breaks=c(3,13,23,33,43),labels=c(-4,-2,0,2,4))+
      scale_y_continuous(breaks=c(1,2,3,4,5,9,13),
                         labels=TeX(c("0","$10^{-4}","$10^{-3}","$10^{-2}","0.1","0.5","0.9")))+
      geom_text(data=sub_num_df,aes(x=10,y=12,label=paste0("n=",n_segre)))+
      labs(title=TeX(p_title),x=TeX("$Effect size (\\omega)"),y="Allele frequency")+
      theme(panel.grid.major = element_line(),legend.position="bottom",aspect.ratio = 0.6,
            axis.text = element_text(size=9))
    
    assign(paste0("Vs",Vs0,"sd",sdMutation0),matrix_p)
  }
}

pdf("output/genetic_architecture_traits.pdf",width=8.2,height=11)
ggarrange(Vs1sd0.05,Vs1sd0.1,Vs1sd0.5,common.legend = T,nrow=1)
ggarrange(Vs10sd0.05,Vs10sd0.1,Vs10sd0.5,common.legend = T,nrow=1)
ggarrange(Vs100sd0.05,Vs100sd0.1,Vs100sd0.5,common.legend = T,nrow=1)
dev.off()
```




disconnect to the database
```{r}
dbDisconnect(singleTrait_db)
```

