---
title: "Summary statistics of the stationary states"
author: "Yuna"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: '2024-01-07'
---

### Working path

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/yuna/B6/simulation/",echo = T,message =  F,warning=F,results = "hide")
```

### Library packages

```{r}
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggpmisc)# add linear regression equation to figures
library(latex2exp)
library(ggpubr)# multiple figures in a page
library(scales)
library(cowplot)# get_legend() & plot_grid() functions
```

### Simulation parameters

```{r}
N=10000
theta=7e-10 * 2e4 * 1e6

paras<-rbind(data.frame(Vs=1,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.35,0.38,0.39,0.4,0.5,0.53,0.55,0.8,1.0,1.2,1.4,1.5,1.6)),
             data.frame(Vs=1,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.0,1.1,1.2,1.3,1.6,1.65,2.2,2.8,2.9,3.0,3.2,3.4,3.6)),
             data.frame(Vs=1,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.0,2.0,3.0,6.5,10.0,11.0,12.0,12.5,13.5,14.0,15.0,16.0,17.0,18.0,18.2,20.0,20.5,21.0)),
             data.frame(Vs=10,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.07,0.08,0.09,0.1,0.11,0.2,0.3,0.4)),
             data.frame(Vs=10,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.15,0.18,0.2,0.22,0.23,0.24,0.3,0.4,0.5,0.6,0.7,0.8)),
             data.frame(Vs=10,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.4,0.5,1.3,2.3,2.32,2.35,2.4,3.3,3.35,4.3,1.0,4.5,4.6,4.9,5.0,5.2,5.4,6.0)),
             data.frame(Vs=100,sdMutation=0.05,speedCoef=c(0.002,0.005,0.01,0.02,0.025,0.03,0.031,0.035,0.04,0.05,0.1,0.2)),
             data.frame(Vs=100,sdMutation=0.1,speedCoef=c(0.002,0.005,0.01,0.02,0.03,0.04,0.045,0.05,0.055,0.057,0.1,0.2,0.3)),
             data.frame(Vs=100,sdMutation=0.5,speedCoef=c(0.002,0.005,0.01,0.02,0.05,0.1,0.2,0.3,0.35,0.38,0.39,0.4,0.5,0.52,0.55,0.7,0.9,1.1,1.3,1.4,1.5))
)
```

### Global plotting settings

```{r}
theme_set(theme_classic())
theme_update(text=element_text(size=13),plot.title=element_text(size=13),
             legend.text=element_text(size=10),legend.title=element_text(size=10,hjust=.5),
             legend.key.size=unit(.3,'cm'),
             legend.key.width=unit(.6,'cm'),
             legend.spacing=unit(0,'cm'),
             legend.background = element_rect(color="grey",fill=NA))

label_sdMutation<-function(string){TeX(paste("$\\omega : $",string))}

negativelog_trans <- function(base = exp(1)) {
    trans <- function(x) log(-x, base)
    inv <- function(x) -base^x
    trans_new(paste0("reverselog_", format(base)), transform = trans, inverse=inv)
}
```


### Stationary state during the last 1000 generations

load data
```{r}
load("data/processed/df2.RData")
```

Stationary state in the stabilizing-selection phase
```{r}
# Lag is stationary
ggplot(SSE_df2,aes(x=generation/N-10,y=gap))+
  geom_line()+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Lag")
```

Stationary state in the moving-optimum phase
```{r}
# Lag is stationary
ggplot(MOE_df2,aes(x=generation/N-10,y=gap))+
  geom_line(aes(group=speedCoef,col=speedCoef*sqrt(SSE_VG)/sdMutation))+
  facet_wrap(sdMutation~Vs,scales="free_y",
             labeller=labeller(Vs=label_both,sdMutation=as_labeller(label_sdMutation,label_parsed)))+
  scale_color_viridis_c()+
  labs(x=TeX("$Generation\\ (\\times 10^4 )"),y="Lag",col = "Speed")
```

Take the average of the 1000 generations
```{r}
# take average
df<-group_by(MOE_df2,Vs,sdMutation,speedCoef)%>%
  summarise(MOE_VG=mean(varPheno_mean),
            MOE_VG_sd=sd(varPheno_mean),
            MOE_gap=mean(gap),
            MOE_gap_sd=sd(gap),
            MOE_fitness=mean(meanFitness_mean))
df<-left_join(df,as.data.frame(parameters),by=c("Vs","sdMutation"))
df<-mutate(df,
           speed=speedCoef * sqrt(SSE_VG),
           sigmaOmega2=sdMutation^2/Vs,
           group=paste0("Vs=",Vs,", $\\sigma_{m}=$",sdMutation))
df<-as.data.frame(df)

so=unique(df$sigmaOmega2)%>%sort
so=as.character(format(so,scientific = T))%>%unique
df$sigmaOmega2<-format(df$sigmaOmega2,scientific=T)
df$sigmaOmega2<-factor(df$sigmaOmega2,level=so,labels=so)
```


### Critical speed
```{r}
#critical speed at specific fitness cutoff
fit0.1<-filter(df,MOE_fitness>0.1)%>%group_by(Vs,sdMutation)%>%slice_max(speedCoef)%>%mutate(Cutoff="0.1")
fit0.01<-filter(df,MOE_fitness>0.01)%>%group_by(Vs,sdMutation)%>%slice_max(speedCoef)%>%mutate(Cutoff="0.01")
fit0.001<-filter(df,MOE_fitness>0.001)%>%group_by(Vs,sdMutation)%>%slice_max(speedCoef)%>%mutate(Cutoff="0.001")
fit0<-group_by(df,Vs,sdMutation)%>%slice_max(speedCoef)%>%mutate(Cutoff="0")

fit_df<-rbind(fit0.1,fit0.01,fit0.001)
# if simulate more, fit0.1$speedCoef<-c(0.38,1.15,12.2,0.072,0.185,2.35,0.022,0.042,0.37)

# generate equations for the power-law
equation_label<-split(fit_df,fit_df$Cutoff)%>%
  lapply(FUN = function(subdf){
    x=log10(subdf$sdMutation^2/subdf$Vs)
    y=log10(subdf$speed/subdf$sdMutation/theta)
    lm_model<-lm(y~x)
    
    a=10^(lm_model$coefficients[1])
    b=lm_model$coefficients[2]
    label=sprintf("$%.2f \\chi^{%.2f}$",a,b)
    
    return(label)
  })
equation_label<-data.frame(Cutoff=names(equation_label),label=unlist(equation_label))
equation_label<-mutate(equation_label,x_pos=1e-4,y_pos=c(0.1,0.08,0.06))

vcrit_p<-ggplot(fit_df,aes(x=sdMutation^2/Vs,y=speed/sdMutation/theta,col=Cutoff))+
  geom_point()+
  stat_poly_line()+
  # stat_poly_eq(use_label(c("eq", "R2")))+
  geom_text(data = equation_label,
    aes(x = x_pos, y = y_pos,label = TeX(label, output = "character")),
    parse = TRUE,show.legend = F)+
  scale_x_continuous(trans='log10',labels = trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',breaks = c(0.01, 0.1),labels = trans_format("log10", math_format(10^.x)))+
  scale_color_manual(values=c("#004488","#DDAA33","#bb5566"),breaks=c("0.001","0.01","0.1"),
                     labels=unname(TeX(c("$w_{th}=0.001","$w_{th}=0.01","$w_{th}=0.1"))))+
  labs(x=TeX("$\\chi"),y=TeX("$Threshold\\ speed\\ (\\omega\\theta)"),col="Fitness threshold")+
  theme(legend.position= c(0.7,0.2))+
  annotation_logticks(sides = "bl")

# speeds in units of sigma_g
equation_label_sigmaG<-split(fit_df,fit_df$Cutoff)%>%
  lapply(FUN = function(subdf){
    x=log10(subdf$sdMutation^2/subdf$Vs)
    y=log10(subdf$speedCoef)
    lm_model<-lm(y~x)
    
    a=10^(lm_model$coefficients[1])
    b=lm_model$coefficients[2]
    label=sprintf("$%.2f \\chi^{%.2f}$",a,b)
    
    return(label)
  })
equation_label_sigmaG<-data.frame(Cutoff=names(equation_label_sigmaG),label=unlist(equation_label_sigmaG))
equation_label_sigmaG<-mutate(equation_label_sigmaG,x_pos=1e-4,y_pos=c(10,6,1))

vcrit_p_sigmaG<-ggplot(fit_df,aes(x=sdMutation^2/Vs,y=speedCoef,col=Cutoff))+
  geom_point()+
  stat_poly_line()+
  # stat_poly_eq(use_label(c("eq", "R2")))+
  geom_text(data = equation_label_sigmaG,
    aes(x = x_pos, y = y_pos,label = TeX(label, output = "character")),
    parse = TRUE,show.legend = F)+
  scale_x_continuous(trans='log10',labels = trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels = trans_format("log10", math_format(10^.x)))+
  scale_color_manual(values=c("#004488","#DDAA33","#bb5566"),breaks=c("0.001","0.01","0.1"),
                     labels=unname(TeX(c("$w_{th}=0.001","$w_{th}=0.01","$w_{th}=0.1"))))+
  labs(x=TeX("$\\chi"),y=TeX("$Threshold\\ speed\\ (\\sigma_{G(S)})"),col="Fitness threshold")+
  # geom_hline(yintercept = c(0.04125416,0.79),linetype="dashed")+ # first one is from Maize GE trait, the 2nd one might be from Darwin Finches
  theme(legend.position= c(0.7,0.2))+
  annotation_logticks(sides = "bl")

vcrit_p
vcrit_p_sigmaG
plot(fit0.1$speedCoef,fit0$speedCoef)
abline(0,1)

# pdf("output/threshold_speed_polygenic.pdf", width=4,height=3)
# vcrit_p
# dev.off()
# pdf("output/threshold_speed_polygenic_scaledByVar.pdf",weight=4,height=3)
# vcrit_p_sigmaG
# dev.off()
```


### Lag ~ shift speed

```{r}
ggplot(df,aes(x=speed,y=MOE_gap,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x="Shift speed",y="Lag")
```

Scaling by omega
```{r}
lag_p<-ggplot(df,aes(x=speed/sdMutation/theta,y=MOE_gap/sdMutation,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ lag\\ (\\omega)"))

lag_p+theme(legend.position = "inside",legend.position.inside = c(0.2,0.6))

# pdf("output/lag_polygenic.pdf",width=4,height=3)
# lag_p+theme(legend.position = "inside",legend.position.inside = c(0.2,0.6))
# dev.off()
```

Scaling by sigma_g
```{r}
lag_p2<-ggplot(df,aes(x=speed/sqrt(SSE_VG),y=MOE_gap,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$Lag"))

lag_p2
```

### Fitness ~ shift speed

```{r}
ggplot(df,aes(x=speed,y=MOE_fitness,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x="Shift speed",y='Fitness')
```

Scaling by omega
```{r}
fit_p<-ggplot(df,aes(x=speed/sdMutation/theta,y=MOE_fitness,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y="Mean fitness")

fit_p+theme(legend.position = "inside",legend.position.inside = c(0.2,0.4))

# pdf("fitness_polygenic.pdf",width=4,height=3)
# fit_p+theme(legend.position = "inside",legend.position.inside = c(0.2,0.4))
# dev.off()
```

Scaling by sigma_g
```{r}
fit_p2<-ggplot(df,aes(x=speed/sqrt(SSE_VG),y=MOE_fitness,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y="Mean fitness")

fit_p2
```

### Genetic variance ~ shift speed

```{r}
ggplot(df,aes(x=speed,y=MOE_VG,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x="Shift speed",y="Genetic variance")
```

Scaling by omega
```{r}
vp_p<-ggplot(df,aes(x=speed/sdMutation/theta,y=MOE_VG/sdMutation^2,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$\\Genetic\\ variance\\ (\\omega^{2})"))

vp_p+theme(legend.position = "inside",legend.position.inside = c(0.8,0.4))

# pdf("Vg_polygenic.pdf",width=4,height=3)
# vp_p+theme(legend.position = "inside",legend.position.inside = c(0.8,0.3),
#            legend.text=element_text(size=8),legend.title=element_text(size=8,hjust=.5),
#            legend.key.size=unit(.2,'cm'),legend.key.width=unit(.4,'cm'))
# dev.off()
```

Scaling by sigma_g
```{r}
vp_p2<-ggplot(df,aes(x=speed/sqrt(SSE_VG),y=MOE_VG,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$\\Genetic\\ variance"))

vp_p2
```



# Summarize mutations

## Summarize fixations

load data
```{r}
load("data/processed/muta2.RData")
```

Take average of every generation at the last 1000 generations and the average of 100 replicates
```{r}
moe_fixed_mean<-group_by(moe_fixed,Vs,sdMutation,speedCoef)%>%
  summarise_at(vars(-replicate),mean,na.rm=T)
moe_fixed_sd<-group_by(moe_fixed,Vs,sdMutation,speedCoef)%>%
  summarise_at(vars(-replicate),sd,na.rm=T)
colnames(moe_fixed_sd)[-c(1:3)]<-paste0(colnames(moe_fixed_sd)[-c(1:3)],"_sd")

moe_fixed2<-left_join(moe_fixed_mean,moe_fixed_sd,by=c("Vs","sdMutation","speedCoef"))

moe_fixed2<-left_join(moe_fixed2,parameters,by=c("Vs","sdMutation"))
moe_fixed2<-mutate(moe_fixed2,speed=speedCoef*sqrt(SSE_VG),
                   sigmaOmega2=sdMutation^2/Vs,group=paste0("Vs",Vs,"sd",sdMutation))

so=unique(moe_fixed2$sigmaOmega2)%>%sort
so=as.character(format(so,scientific = T))%>%unique
moe_fixed2$sigmaOmega2<-format(moe_fixed2$sigmaOmega2,scientific=T)
moe_fixed2$sigmaOmega2<-factor(moe_fixed2$sigmaOmega2,level=so,label=so)
```

### Lag ~ fixed effect
```{r}
df_moe_fixed2<-left_join(df,moe_fixed2[,1:(ncol(moe_fixed2)-4)],by=c("Vs","sdMutation","speedCoef"))

lag_effect_p<-ggplot(df_moe_fixed2,aes(x=MOE_gap/sdMutation,y=mean_effect/sdMutation,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10')+
  labs(x=TeX("$Lag\\ (\\omega)"),y=TeX("$Mean\\ fixed\\ effect\\ size\\ (\\omega)"))

lag_effect_p2<-ggplot(df_moe_fixed2,aes(x=MOE_gap,y=mean_effect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10')+
  labs(x="Lag",y="Mean fixed effect size")

ggarrange(lag_effect_p,lag_effect_p2,ncol=1)

```


### Fixation number ~ speed

```{r}
ggplot(moe_fixed2,aes(x=speed,y=n_fixed/1000,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x="Shift speed",y="Number of fixations")

# non-log x axis
ggplot(moe_fixed2,aes(x=speed,y=n_fixed/1000,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  labs(x="Shift speed",y="Number of fixations")
```

Scaling by omega
```{r}
fn_p<-ggplot(moe_fixed2,aes(x=speed/sdMutation/theta,y=n_fixed/1000,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y="Number of fixations")

fn_p

# non-log x axis
fn_p_nonlog<-ggplot(moe_fixed2,aes(x=speed/sdMutation/theta,y=n_fixed/1000))+
  stat_poly_line(col="red",linetype="dashed",se=F)+
  stat_poly_eq(use_label(c("eq", "R2")))+
  geom_point(aes(col=sigmaOmega2))+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y="Number of fixations")

fn_p_nonlog

# pdf("output/fixation_number_polygenic.pdf",width=4,height=3)
# fn_p_nonlog+theme(legend.position = "inside",legend.position.inside = c(0.8,0.3))
# dev.off()
```

Scaling by sigma_g
```{r}
fn_p2<-ggplot(moe_fixed2,aes(x=speed/sqrt(SSE_VG),y=n_fixed/1000,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y="Number of fixations")

fn_p2

# non-log x axis
ggplot(moe_fixed2,aes(x=speed/sqrt(SSE_VG),y=n_fixed/1000,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y="Number of fixations")

```

### Fixation time ~ speed

```{r}
ggplot(moe_fixed2,aes(x=speed,y=mean_fixedtime,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x="Shift speed",y="Fixation time")
```

Scaling by omega
```{r}
ft_p<-ggplot(moe_fixed2,aes(x=speed/sdMutation/theta,y=mean_fixedtime,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y="Fixation time")

ft_p
```

Scaling by sigma_g
```{r}
ft_p2<-ggplot(moe_fixed2,aes(x=speed/sqrt(SSE_VG),y=mean_fixedtime,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y="Fixation time")

ft_p2
```

### Mean fixed effect size ~ speed

```{r}
ggplot(moe_fixed2,aes(x=speed,y=mean_effect,group=group,col=sigmaOmega2))+
  geom_hline(aes(yintercept=sdMutation * sqrt(2/pi)))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  annotate("text",x=0.8,y=0.05 * sqrt(2/pi)+0.005,label=TeX("$\\omega=0.05",output="character"),parse=T)+
  annotate("text",x=0.8,y=0.1 * sqrt(2/pi)+0.02,label=TeX("$\\omega=0.1",output="character"),parse=T)+
  annotate("text",x=0.8,y=0.5 * sqrt(2/pi)+0.05,label=TeX("$\\omega=0.5",output="character"),parse=T)+
  labs(x="Shift speed",y=TeX("$Mean\\ \\alpha\\ of\\ fixations"))
```

Scaling by omega
```{r}
mes_p<-ggplot(moe_fixed2,aes(x=speed/sdMutation/theta,y=mean_effect/sdMutation,group=group,col=sigmaOmega2))+
  # geom_hline(yintercept=sqrt(2/pi))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  #scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  #scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ \\alpha\\ of\\ fixations\\ (\\omega)"))

mes_p

pdf("fixation_effect_polygenic.pdf",width=4,height=3)
mes_p+theme(legend.position = "inside",legend.position.inside = c(0.8,0.3))
dev.off()
```

Scaling by sigma_g
```{r}
mes_p2<-ggplot(moe_fixed2,aes(x=speed/sqrt(SSE_VG),y=mean_effect,group=group,col=sigmaOmega2))+
  # geom_hline(yintercept=sqrt(2/pi))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$Mean\\ \\alpha\\ of\\ fixations"))

mes_p2
```

### Mean standard deviation of fixed effect size \~ speed

```{r}
ggplot(moe_fixed2,aes(x=speed,y=sd_effect,group=group,col=sigmaOmega2))+
  geom_hline(aes(yintercept=sdMutation * sqrt(1-2/pi)))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  annotate("text",x=0.8,y=0.05 * sqrt(1-2/pi)+0.005,label=TeX("$\\omega=0.05",output="character"),parse=T)+
  annotate("text",x=0.8,y=0.1 * sqrt(1-2/pi)+0.01,label=TeX("$\\omega=0.1",output="character"),parse=T)+
  annotate("text",x=0.8,y=0.5 * sqrt(1-2/pi)+0.02,label=TeX("$\\omega=0.5",output="character"),parse=T)+
  labs(x="Shift speed",y=TeX("$Sd(\\alpha)\\ of\\ fixations"))
```

Scaling by omega
```{r}
sde_p<-ggplot(moe_fixed2,aes(x=speed/sdMutation/theta,y=sd_effect/sdMutation,group=group,col=sigmaOmega2))+
  # geom_hline(yintercept=sqrt(1-2/pi))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Sd(\\alpha)\\ of\\ fixations\\ (\\omega)"))   

sde_p
```

Scaling by sigma_g
```{r}
sde_p2<-ggplot(moe_fixed2,aes(x=speed/sqrt(SSE_VG),y=sd_effect,group=group,col=sigmaOmega2))+
  # geom_hline(yintercept=sqrt(1-2/pi))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$Sd(\\alpha)\\ of\\ fixations"))
       
sde_p2
```


### Mean positive fixed effect \~ speed

```{r}
ggplot(moe_fixed2,aes(x=speed,y=mean_posEffect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x="Shift speed",y="Mean positive fixed effect")
```

Scaling by omega
```{r}
ggplot(moe_fixed2,aes(x=speed/sdMutation/theta,y=mean_posEffect/sdMutation,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ positive\\ fixed\\ effect\\ (\\omega)"))
```

### Mean negative fixed effect \~ speed

```{r}
ggplot(moe_fixed2,aes(x=speed,y=mean_negEffect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans=negativelog_trans(10),breaks=c(0,-0.1,-0.2,-0.4,-0.6))+
  labs(x="Shift speed",y="Mean negative fixed effect")
```

Scaling by omega
```{r}
ggplot(moe_fixed2,aes(x=speed/sdMutation/theta,y=mean_negEffect/sdMutation,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans=negativelog_trans(10),breaks=c(0,-0.25,-0.75,-1))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ negative\\ fixed\\ effect\\ (\\omega)"))
```


## Summarize segregating mutations

Take average of every generation at the last 1000 generations and the average of 100 replicates
```{r}
moe_muta_mean<-subset(shift_muta2,generation>19.9*N)%>%
  group_by(Vs,sdMutation,speedCoef,replicate)%>%
  summarise_at(vars(-generation),mean,na.rm=T)%>%
  group_by(Vs,sdMutation,speedCoef)%>%summarise_at(vars(-replicate),mean,na.rm=T)
moe_muta_sd<-subset(shift_muta2,generation>19.9*N)%>%
  group_by(Vs,sdMutation,speedCoef,replicate)%>%
  summarise_at(vars(-generation),mean,na.rm=T)%>%
  group_by(Vs,sdMutation,speedCoef)%>%summarise_at(vars(-replicate),sd,na.rm=T)
colnames(moe_muta_sd)[-c(1:3)]<-paste0(colnames(moe_muta_sd)[-c(1:3)],'_sd')
moe_muta2<-left_join(moe_muta_mean,moe_muta_sd,by=c('Vs','sdMutation','speedCoef'))

moe_muta2<-left_join(moe_muta2,parameters,by=c('Vs','sdMutation'))
moe_muta2<-mutate(moe_muta2,speed=speedCoef * sqrt(SSE_VG),
                   sigmaOmega2=sdMutation^2/Vs,
                   group=paste0("Vs=",Vs,", $\\sigma_{m}=$",sdMutation))

so=unique(moe_muta2$sigmaOmega2)%>%sort
so=as.character(format(so,scientific = T))%>%unique
moe_muta2$sigmaOmega2<-format(moe_muta2$sigmaOmega2,scientific=T)
moe_muta2$sigmaOmega2<-factor(moe_muta2$sigmaOmega2,level=so,label=so)
```

### Segregating number ~ speed

```{r}
ggplot(moe_muta2,aes(x=speed,y=n_segregations,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x="Shift speed",y="Number of segregating sites")
```

Scaling by omega
```{r}
sn_p<-ggplot(moe_muta2,aes(x=speed/sdMutation/theta,y=n_segregations,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y="Number of segregating sites")

sn_p

pdf("segregating_number_polygenic.pdf",width=4,height=3)
sn_p+theme(legend.position = "inside",legend.position.inside = c(0.4,0.3),
           legend.text=element_text(size=8),legend.title=element_text(size=8,hjust=.5),
           legend.key.size=unit(.2,'cm'),legend.key.width=unit(.4,'cm'))
dev.off()
```

Scaling by sigma_g
```{r}
sn_p2<-ggplot(moe_muta2,aes(x=speed/sqrt(SSE_VG),y=n_segregations,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y="Number of segregating sites")

sn_p2
```

### Proportion of positive segregating mutations ~ speed

```{r}
ggplot(moe_muta2,aes(x=speed,y=n_posSegregations/n_segregations,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(labels=percent)+
  labs(x="Shift speed",y="Proportion of positive segregating mutations")
```

Scaling by omega
```{r}
ggplot(moe_muta2,aes(x=speed/sdMutation/theta,y=n_posSegregations/n_segregations,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(labels=percent)+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y="Proportion of positive segregating mutations")
```

Scaling by sigma_g
```{r}
ggplot(moe_muta2,aes(x=speed/sqrt(SSE_VG),y=n_posSegregations/n_segregations,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(labels=percent)+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y="Proportion of positive segregating mutations")
```

### Mean segregating effect size ~ speed

```{r}
ggplot(moe_muta2,aes(x=speed,y=mean_effect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x="Shift speed",y=TeX("$Mean\\ \\alpha\\ of\\ segregating\\ sites"))
```

Scaling by omega
```{r}
mse_p<-ggplot(moe_muta2,aes(x=speed/sdMutation/theta,y=mean_effect/sdMutation,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ \\alpha\\ of\\ segregating\\ sites\\ (\\omega)"))

mse_p

pdf("segregating_effect_polygenic.pdf",width=4,height=3)
mse_p+theme(legend.position = "inside",legend.position.inside = c(0.8,0.35),
           legend.text=element_text(size=8),legend.title=element_text(size=8,hjust=.5),
           legend.key.size=unit(.2,'cm'),legend.key.width=unit(.4,'cm'))
dev.off()
```

Scaling by sigma_g
```{r}
mse_p2<-ggplot(moe_muta2,aes(x=speed/sqrt(SSE_VG),y=mean_effect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$Mean\\ \\alpha\\ of\\ segregating\\ sites"))

mse_p2
```

### Mean positive segregating effect ~ speed

```{r}
ggplot(moe_muta2,aes(x=speed,y=mean_posEffect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x="Shift speed",y="Mean positive segregating effect")
```

Scaled by omega
```{r}
ggplot(moe_muta2,aes(x=speed/sdMutation/theta,y=mean_posEffect/sdMutation,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ positive\\ segregating\\ effect\\ (\\omega)"))
```

Scaled by sigma_g
```{r}
ggplot(moe_muta2,aes(x=speed/sqrt(SSE_VG),y=mean_posEffect,group=group,col=sigmaOmega2))+
  geom_point()+
 geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans='log10')+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$Mean\\ positive\\ segregating\\ effect"))
```

### Mean negative effect ~ speed

```{r}
ggplot(moe_muta2,aes(x=speed,y=mean_negEffect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans=negativelog_trans(10))+
  labs(x="Shift speed",y="Mean negative segregating effect")
```

Scaled by omega
```{r}
ggplot(moe_muta2,aes(x=speed/sdMutation/theta,y=mean_negEffect/sdMutation,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans=negativelog_trans(10))+
  labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ negative\\ segregating\\ effect\\ (\\omega)"))
```

Scaled by sigma_g
```{r}
ggplot(moe_muta2,aes(x=speed/sqrt(SSE_VG),y=mean_negEffect,group=group,col=sigmaOmega2))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+
  scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+
  scale_y_continuous(trans=negativelog_trans(10))+
  labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y="Mean negative segregating effect")
```


<!-- ### Weighted mean segregating effect size ~ speed -->

<!-- ```{r} -->
<!-- load("summ_weighted.RData") -->

<!-- moe_muta_weightedMeanSd<-left_join(moe_muta_weightedMeanSd,parameters,by=c('Vs','sdMutation')) -->
<!-- moe_muta_weightedMeanSd<-mutate(moe_muta_weightedMeanSd,speed=speedCoef * sqrt(SSE_VG), -->
<!--                    sigmaOmega2=sdMutation^2/Vs, -->
<!--                    group=paste0("Vs=",Vs,", $\\sigma_{m}=$",sdMutation)) -->

<!-- so=unique(moe_muta_weightedMeanSd$sigmaOmega2)%>%sort -->
<!-- so=as.character(format(so,scientific = T))%>%unique -->
<!-- moe_muta_weightedMeanSd$sigmaOmega2<-format(moe_muta_weightedMeanSd$sigmaOmega2,scientific=T) -->
<!-- moe_muta_weightedMeanSd$sigmaOmega2<-factor(moe_muta_weightedMeanSd$sigmaOmega2,level=so,label=so) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ggplot(moe_muta_weightedMeanSd,aes(x=speed,y=mean_effect,group=group,col=sigmaOmega2))+ -->
<!--   geom_point()+ -->
<!--   geom_line()+ -->
<!--   scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+ -->
<!--   scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+ -->
<!--   scale_y_continuous(trans='log10')+ -->
<!--   labs(x="Shift speed",y=TeX("$Mean\\ \\alpha\\ of\\ segregating\\ sites")) -->
<!-- ``` -->

<!-- Scaling by omega -->
<!-- ```{r} -->
<!-- ggplot(moe_muta_weightedMeanSd,aes(x=speed/sdMutation/theta,y=mean_effect/sdMutation,group=group,col=sigmaOmega2))+ -->
<!--   geom_point()+ -->
<!--   geom_line()+ -->
<!--   scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+ -->
<!--   scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+ -->
<!--   scale_y_continuous(trans='log10')+ -->
<!--   labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Mean\\ \\alpha\\ of\\ segregating\\ sites\\ (\\omega)")) -->
<!-- ``` -->

<!-- Scaling by sigma_g -->
<!-- ```{r} -->
<!-- ggplot(moe_muta_weightedMeanSd,aes(x=speed/sqrt(SSE_VG),y=mean_effect,group=group,col=sigmaOmega2))+ -->
<!--   geom_point()+ -->
<!--   geom_line()+ -->
<!--   scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+ -->
<!--   scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+ -->
<!--   scale_y_continuous(trans='log10')+ -->
<!--   labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$Mean\\ \\alpha\\ of\\ segregating\\ sites")) -->
<!-- ``` -->

<!-- ### Weighted standard deviation of segregating effect size ~ speed -->

<!-- ```{r} -->
<!-- ggplot(moe_muta_weightedMeanSd,aes(x=speed,y=sd_effect,group=group,col=sigmaOmega2))+ -->
<!--   geom_point()+ -->
<!--   geom_line()+ -->
<!--   scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+ -->
<!--   scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+ -->
<!--   scale_y_continuous(trans='log10')+ -->
<!--   labs(x="Shift speed",y=TeX("$Sd(\\alpha)\\ of\\ segregating\\ sites")) -->
<!-- ``` -->

<!-- Scaling by omega -->
<!-- ```{r} -->
<!-- ggplot(moe_muta_weightedMeanSd,aes(x=speed/sdMutation/theta,y=sd_effect/sdMutation,group=group,col=sigmaOmega2))+ -->
<!--   geom_point()+ -->
<!--   geom_line()+ -->
<!--   scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+ -->
<!--   scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+ -->
<!--   scale_y_continuous(trans='log10')+ -->
<!--   labs(x=TeX("$Shift\\ speed\\ (\\omega\\theta / generation)"),y=TeX("$Sd(\\alpha)\\ of\\ segregating\\ sites\\ (\\omega)")) -->
<!-- ``` -->

<!-- Scaling by sigma_g -->
<!-- ```{r} -->
<!-- ggplot(moe_muta_weightedMeanSd,aes(x=speed/sqrt(SSE_VG),y=sd_effect,group=group,col=sigmaOmega2))+ -->
<!--   geom_point()+ -->
<!--   geom_line()+ -->
<!--   scale_color_viridis_d(name=TeX("$\\chi"),begin=0,end=0.9)+ -->
<!--   scale_x_continuous(trans='log10',labels=trans_format("log10", math_format(10^.x)))+ -->
<!--   scale_y_continuous(trans='log10')+ -->
<!--   labs(x=TeX("$Shift\\ speed\\ (\\sigma_{G(S)} / generation)"),y=TeX("$Sd(\\alpha)\\ of\\ segregating\\ sites")) -->
<!-- ``` -->

